### 2021-09-17 UOWN_MADA_CC.Rmd
Project - Part 1_CC

---
title: "MADA Project_UOWN"
author: "C. Coleman"
date: "9/16/2021"
output: html_document
---

#Briefly describe what the data is, how it was collected, where you will get (or got) it from. How many observations do you have, what was measured? Anything else important to report about the data?

#The data I am using is a mixture of Upper Oconee Watershed Network (UOWN) publically avaialable data and data available on the UOWN ShinyApp. This data is part of a citizen science training effort for routine watershed-scale monitoring of the upper oconee river watershed. Trained freshwater scientists assist in training members of the community to collect both chemical and biological stream characteristics that are useful in describing stream health. All data is compliled and posted in raw form on the UOWN website. Of interest to me are the following variables: WSID (ID number),	biological_score,	conductivity, turbidity,	po4.mgL,	no3.mgL,	pH,	temperature,	e.coli.cfu,	month,	year,	day. The raw data shows 3,075 observations before cleaning.

#Load import packages:
```{r}
library(tidyverse)
library(readxl)
```

#Lodaing data and providing a summary for the raw data
```{r}
data_location <- here::here("data","raw_data","UOWN_data_master_spring2021 (2).xlsx")

rawdata <- readxl::read_excel(data_location)
```

#Explain the question you want to answer using the data. What will be your outcome(s) of interest (if any)? What (if any) specific predictors will you focus on? What relations/patterns are you looking for in the data?

#When monitoring freshwater streams, stream health is typically measured by its physical charachteristics (chemistry, surrounding land use area, channelication, etc.); Microbial (fecal indicator), and ecology (diversity of indicator macroinvertebrates). However, these thre categories are usually analyzed separately to assess stream health. As such, my question is how well are these parameters correleated with each other. Specifically, can you use one of the indicators mentioned (physical, microbial, and ecological) to describe both of the others to make a statement about overall stream health?

I would expect that because the all three are correlated with urbanization/anthropocentric impact, a negative impact in on indicator category will indicate a correlated negative impact in the other categories. The only caveat is that because level of noise in stream health data, significant correlations may be difficult to find.

#As much as you know, suggest how you will analyze it. At this stage in the course, we haven’t covered analysis approaches yet, so you can keep things vague and non-technical here.

UOWN has provided a great deal of data dating back to 2003. First,I will analyze and compare trends in the change of chemical (conductivity, turbidity,	po4.mgL,	no3.mgL); ecological(Biological Score); and microbial indicator data (e.coli CFU). Generally speaking, Increases in chemical properties like conductivity, turbidity,	po4.mgL,	no3.mgL and microbial indicator abundance (i.e. ecoli CFU) indicate decreased stream health. While a decrease in biological score indicates decrease in stream health. Next, since biological scores are used as a measure of broader ecological health, I will run t-tests on if the other factors significantly effect biological score. Finally, I will test the interaction of all three vie multivariate analysis.

### 2021-09-17 epitweetr_setup.Rmd
Project - Part 1_CC

---
title: "epitweetr"
author: "C. Coleman"
date: "9/13/2021"
output: html_document
---

#*Set-up*
#Checking if Java Home is in the System environment. Make sure it doesn't return "NULL".
```{r}
Sys.getenv("JAVA_HOME")
```

#You have to go online and download Rtools to run epitweetr. Once downloaded, install these packages.
```{r}
install.packages(DT, type = "source")
```

#Load epitweetr.
```{r}
library(epitweetr)
```

#opening epitweetr dashboard and making quere for covid-19 searches.
```{r}
epitweetr::epitweetr_app("data_dir")
```
#*Twitter Authentification*
#Install rtweet package. 
```{r}
install.packages(rtweet, type = "source")
```
#Load rtweet
```{r}
load(rtweet)
```

### 2021-09-17 UOWN_MADA_CC.Rmd
Project Part 1 - CC

@@
 rawdata <- readxl::read_excel(data_location)
+
+summary(rawdata)
 ```
@@
 
-#When monitoring freshwater streams, stream health is typically measured by its physical charachteristics (chemistry, surrounding land use area, channelication, etc.); Microbial (fecal indicator), and ecology (diversity of indicator macroinvertebrates).
+When monitoring freshwater streams, stream health is typically measured by its physical charachteristics (chemistry, surrounding land use area, channelication, etc.); Microbial (fecal indicator), and ecology (diversity of indicator macroinvertebrates).
 However, these thre categories are usually analyzed separately to assess stream health.
@@
 
-I would expect that because the all three are correlated with urbanization/anthropocentric impact, a negative impact in on indicator category will indicate a correlated negative impact in the other categories.
+I would expect that because the all three are correlated with urbanization/anthropocentric impact, a negative impact in one indicator category will indicate a correlated negative impact in the other categories.
 The only caveat is that because level of noise in stream health data, significant correlations may be difficult to find.
@@
 UOWN has provided a great deal of data dating back to 2003.
-First,I will analyze and compare trends in the change of chemical (conductivity, turbidity,	po4.mgL,	no3.mgL); ecological(Biological Score); and microbial indicator data (e.coli CFU).
+First,I will need to wrangle missing data values that are a result of inconsitent sampling efforts.
+This will make the data easier to work with and narrow down a time frame for the data that is consistent and free of sampling bias.
+Next, I will analyze and compare trends in the change of chemical (conductivity, turbidity,	po4.mgL,	no3.mgL); ecological(Biological Score); and microbial indicator data (e.coli CFU).
 Generally speaking, Increases in chemical properties like conductivity, turbidity,	po4.mgL,	no3.mgL and microbial indicator abundance (i.e.

### 2021-10-08 README.Rmd
Project Part 2 - CC

--- 2021-09-17 UOWN_MADA_CC.Rmd
@@
 
-#Briefly describe what the data is, how it was collected, where you will get (or got) it from. How many observations do you have, what was measured? Anything else important to report about the data?
+# Overview: Comparison of Stream Health indicators
 
-#The data I am using is a mixture of Upper Oconee Watershed Network (UOWN) publically avaialable data and data available on the UOWN ShinyApp.
+### The data I am using is the publically avaialable data provided by the Upper Oconee Watershed Network (UOWN).
 This data is part of a citizen science training effort for routine watershed-scale monitoring of the upper oconee river watershed.
@@
 The raw data shows 3,075 observations before cleaning.
+However, while cleaning I had to eliminate "po4.mgl" and "temperature" du to significant amounts of missing data.
 
-#Load import packages:
+### For quick reference, here is hoe to load the raw data:
+
+#### Load import packages:
 ```{r}
@@
 
-#Lodaing data and providing a summary for the raw data
+#### Lodaing data and providing a summary for the raw data
 ```{r}
@@
 
-summary(rawdata)
+glimpse(data_cleaning)
 ```
 
-#Explain the question you want to answer using the data.
-What will be your outcome(s) of interest (if any)?
-What (if any) specific predictors will you focus on?
-What relations/patterns are you looking for in the data?
+### Question:
 
-When monitoring freshwater streams, stream health is typically measured by its physical charachteristics (chemistry, surrounding land use area, channelication, etc.); Microbial (fecal indicator), and ecology (diversity of indicator macroinvertebrates).
-However, these thre categories are usually analyzed separately to assess stream health.
-As such, my question is how well are these parameters correleated with each other.
+#### When monitoring freshwater streams, stream health is typically measured by its physical characteristics (chemistry, surrounding land use area, channelization, etc.); Microbial (fecal indicator), and ecology (diversity of indicator macroinvertebrates).
+However, these categories are usually analyzed separately to assess stream health.
+As such, my question is how well are these parameters correlated with each other.
 Specifically, can you use one of the indicators mentioned (physical, microbial, and ecological) to describe both of the others to make a statement about overall stream health?
@@
 UOWN has provided a great deal of data dating back to 2003.
-First,I will need to wrangle missing data values that are a result of inconsitent sampling efforts.
+First,I will need to wrangle missing data values that are a result of inconsistent sampling efforts.
 This will make the data easier to work with and narrow down a time frame for the data that is consistent and free of sampling bias.
 Next, I will analyze and compare trends in the change of chemical (conductivity, turbidity,	po4.mgL,	no3.mgL); ecological(Biological Score); and microbial indicator data (e.coli CFU).
-Generally speaking, Increases in chemical properties like conductivity, turbidity,	po4.mgL,	no3.mgL and microbial indicator abundance (i.e.
+Generally speaking, increases in chemical properties like conductivity, turbidity,	po4.mgL,	no3.mgL and microbial indicator abundance (i.e.
 ecoli CFU) indicate decreased stream health.
@@
 Next, since biological scores are used as a measure of broader ecological health, I will run t-tests on if the other factors significantly effect biological score.
-Finally, I will test the interaction of all three vie multivariate analysis.
+Finally, I will test the interaction of all three via multivariate analysis.

### 2021-10-08 code/processing_code/Data_Exploration.Rmd
Project Part 2 - CC

---
title: "Data Exploration"
author: "C. Coleman"
date: "10/7/2021"
output: word_document
---

# The following is an exploration of Stream Health Indicator correlation and comparison. The exploration will be structured as follows:
### Each dataframe (MIDO, NORO, and BICO) will be analyzed separately. Within each dataframe, there are data explaining biological indicator values, stream chemistry, and fecal indicators over time. The only group with multiple components is stream chemistry. I plan to do a series of correlative analysis showing if there is some sort of relationship between the three indicator types to help inform further analysis via principle component, multivariate analysis, etc. Therfore, the goal of this exploration is basically to summarize how these indicators correlate.

</br>

#### First, we need to load all the libraries we will need.
```{r}
library(dplyr) 
library(here)
library(tidyverse)
library(tidyr)
library(readr)
library(stringr)
```

#### Loading the data from the processing script. Note that because you are loading a library at the beginning, it may either take a while or take several tries.
```{r}
MIDO <- readRDS("../../data/processed_data/MIDO.rds")
NORO <- readRDS("../../data/processed_data/NORO.rds")
BICO <- readRDS("../../data/processed_data/BICO.rds")
```
 <- source("processingscript.R", local = knitr::knit_global())

#### take a look at the data frames to make sure everything worked correctly.
```{r}
view(MIDO)
view(NORO)
view(BICO)
```

</br>

## MIDO

</br>

### *Biological Indicator:*

#### Biological Vs Conductivity
```{r}
BS_Con <- MIDO %>% 
  ggplot(aes(x=(conductivity.uscm), 
             y=(biological_score))) +
  geom_point(color = "red") +
  geom_smooth(method=lm, color = "red") +
  scale_x_continuous(trans = 'log10') +
  scale_y_continuous(trans = 'log10')

print(BS_Con)
```

#### Now run linear regression for BS_Con.
```{r}
BS_Con_lm <- lm(biological_score ~ conductivity.uscm, data=MIDO)

summary(BS_Con_lm)
```

#### With *m = -0.017* and *p = 0.49*, there is a *negative*, *non-significant* correlation between *biological score* and *conductivity*.

</br>

#### Biological Vs Turbidity
```{r}
BS_Turb <- MIDO %>%
  ggplot(aes(x=turbidity.ntu, y=biological_score)) +
  geom_point(color = "blue") +
   geom_smooth(method=lm, color = "blue") +
  scale_x_continuous(trans = 'log10') +
  scale_y_continuous(trans = 'log10')

print(BS_Turb)
```

#### Now run linear regression for BS_Turb.
```{r}
BS_Turb_lm <- lm(biological_score ~ turbidity.ntu, data=MIDO)

summary(BS_Turb_lm)
```
#### With *m = 0.11* and *p = 0.13*, there is a *positive*, *non-significant* correlation between *biological score* and *turbidity*.

</br>

#### Biological Vs Nitrate Concentration
```{r}
BS_no3 <- MIDO %>%
  ggplot(aes(x=no3.mgL, y=biological_score)) +
  geom_point(color = "green") +
   geom_smooth(method=lm, color = "green") +
  scale_x_continuous(trans = 'log10') +
  scale_y_continuous(trans = 'log10')

print(BS_no3)
```

#### Now run linear regression for BS_no3.
```{r}
BS_no3_lm <- lm(biological_score ~ no3.mgL, data=MIDO)

summary(BS_no3_lm)
```

#### With *m = 0.11* and *p = 0.27*, there is a *positive*, *non-significant* correlation between *biological score* and *no3*.

</br>

#### Biological Vs E. Coli CFU
```{r}
BS_ECFU <- MIDO %>%
  ggplot(aes(x=e.coli.cfu, y=biological_score)) +
  geom_point(color = "purple") +
   geom_smooth(method=lm, color = "purple") +
  scale_x_continuous(trans = 'log10') +
  scale_y_continuous(trans = 'log10')

print(BS_ECFU)
```

#### Now run linear regression for BS_ECFU.
```{r}
BS_ECFU_lm <- lm(biological_score ~ e.coli.cfu, data=MIDO)

summary(BS_ECFU_lm)
```
#### With *m = -0.0007* and *p = 0.39*, there really isn't a relationship at all between *biological score* and *ECFU*.

</br>

### *E. Coli*
</br>
#### E. Coli Vs Conductivity
```{r}
ECFU_Con <- MIDO %>%
  ggplot(aes(x=conductivity.uscm, y=e.coli.cfu)) +
  geom_point(color = "orange") +
   geom_smooth(method=lm, color = "orange") +
  scale_x_continuous(trans = 'log10') +
  scale_y_continuous(trans = 'log10')

print(ECFU_Con)
```

#### Now run linear regression for ECFU_Con.
```{r}
ECFU_Con_lm <- lm(e.coli.cfu ~ conductivity.uscm, data=MIDO)

summary(ECFU_Con_lm)
```
#### With *m = -3.635* and *p = 0.28*, there is a *negative*, *non-significant* correlation between *ECFU* and *Conductivity*.

</br>

#### E. Coli Vs Turbidity
```{r}
ECFU_Turb <- MIDO %>%
  ggplot(aes(x=turbidity.ntu, y=e.coli.cfu)) +
  geom_point(color = "black") +
   geom_smooth(method=lm, color = "black") +
  scale_x_continuous(trans = 'log10') +
  scale_y_continuous(trans = 'log10')

print(ECFU_Turb)
```

#### Now run linear regression for ECFU_Turb.
```{r}
ECFU_Turb_lm <- lm(e.coli.cfu ~ turbidity.ntu, data=MIDO)

summary(ECFU_Turb_lm)
```
#### With *m = 18.6* and *p = 0.07*, there is a *positive*, *non-significant* correlation between *ECFU* and *turbidity*.

</br>

#### E. Coli Vs Nitrate Concentration
```{r}
ECFU_no3 <- MIDO %>%
  ggplot(aes(x=no3.mgL, y=e.coli.cfu)) +
  geom_point(color = "black") +
   geom_smooth(method=lm, color = "black") +
  scale_x_continuous(trans = 'log10') +
  scale_y_continuous(trans = 'log10')

print(ECFU_no3)
```

#### Now run linear regression for ECFU_no3.
```{r}
ECFU_no3_lm <- lm(e.coli.cfu ~ no3.mgL, data=MIDO)

summary(ECFU_no3_lm)
```

#### With *m = -10.61* and *p = 0.60*, there is a *negative*, *non-significant* correlation between *ECFU* and *no3*.

</br>

## NORO

</br>

### *Biological Indicator:*

#### Biological Vs Conductivity
```{r}
N_BS_Con <- NORO %>%
  ggplot(aes(x=conductivity.uscm, y=biological_score)) +
  geom_point(color = "red") +
   geom_smooth(method=lm, color = "red") +
  scale_x_continuous(trans = 'log10') +
  scale_y_continuous(trans = 'log10')

print(N_BS_Con)
```

#### Now run linear regression for N_BS_Con.
```{r}
N_BS_Con_lm <- lm(biological_score ~ conductivity.uscm, data=NORO)

summary(N_BS_Con_lm)
```

#### With *m = -0.023* and *p = 0.0004*, there is a *negative*, *significant* correlation between *biological score* and *conductivity*.

</br>

#### Biological Vs Turbidity
```{r}
N_BS_Turb <- NORO %>%
  ggplot(aes(x=turbidity.ntu, y=biological_score)) +
  geom_point(color = "blue") +
   geom_smooth(method=lm, color = "blue") +
  scale_x_continuous(trans = 'log10') +
  scale_y_continuous(trans = 'log10')

print(N_BS_Turb)
```

#### Now run linear regression for N_BS_Turb.
```{r}
N_BS_Turb_lm <- lm(biological_score ~ turbidity.ntu, data=NORO)

summary(N_BS_Turb_lm)
```

#### With *m = 0.17* and *p = 0.45*, there is a *positive*, *non-significant* correlation between *bilogical score* and *turbidity*.

</br>

#### Biological Vs Nitrate Concentration
```{r}
N_BS_no3 <- NORO %>%
  ggplot(aes(x=no3.mgL, y=biological_score)) +
  geom_point(color = "green") +
   geom_smooth(method=lm, color = "green") +
  scale_x_continuous(trans = 'log10') +
  scale_y_continuous(trans = 'log10')

print(N_BS_no3)
```

#### Now run linear regression for N_BS_no3.
```{r}
N_BS_no3_lm <- lm(biological_score ~ no3.mgL, data=NORO)

summary(N_BS_no3_lm)
```

#### With *m = -0.56* and *p = 0.08*, there is a *negative*, *non-significant* correlation between *biological score* and *no3*.

</br>

#### Biological Vs E. Coli CFU
```{r}
N_BS_ECFU <- NORO %>%
  ggplot(aes(x=e.coli.cfu, y=biological_score)) +
  geom_point(color = "purple") +
   geom_smooth(method=lm, color = "purple") +
  scale_x_continuous(trans = 'log10') +
  scale_y_continuous(trans = 'log10')

print(N_BS_ECFU)
```

#### Now run linear regression for N_BS_ECFU.
```{r}
N_BS_ECFU_lm <- lm(biological_score ~ e.coli.cfu, data=NORO)

summary(N_BS_ECFU_lm)
```

#### With *m = 7.274e-05* and *p = 0.69*, there is not relationship between *Bbiological score* and *ECFU*.

</br>

### *E. Coli*
</br>
#### E. Coli Vs Conductivity
```{r}
N_ECFU_Con <- NORO %>%
  ggplot(aes(x=conductivity.uscm, y=e.coli.cfu)) +
  geom_point(color = "orange") +
   geom_smooth(method=lm, color = "orange") +
  scale_x_continuous(trans = 'log10') +
  scale_y_continuous(trans = 'log10')

print(N_ECFU_Con)
```

#### Now run linear regression for N_ECFU_Con.
```{r}
N_ECFU_Con_lm <- lm(e.coli.cfu ~ conductivity.uscm, data=NORO)

summary(N_ECFU_Con_lm)
```

#### With *m = -8.632* and *p = 0.20*, there is a *negative*, *non-significant* correlation between *ECFU* and *conductivity*.

</br>

#### E. Coli Vs Turbidity
```{r}
N_ECFU_Turb <- NORO %>%
  ggplot(aes(x=turbidity.ntu, y=e.coli.cfu)) +
  geom_point(color = "black") +
   geom_smooth(method=lm, color = "black")

print(N_ECFU_Turb)
```

#### Now run linear regression for N_ECFU_Turb.
```{r}
N_ECFU_Turb_lm <- lm(e.coli.cfu ~ turbidity.ntu, data=NORO)

summary(N_ECFU_Turb_lm)
```

#### With *m = 490* and *p = 0.02*, there is a *positive*, *significant* correlation between *ECFU* and *Turbidity*.

</br>

#### E. Coli Vs Nitrate Concentration
```{r}
N_ECFU_no3 <- NORO %>%
  ggplot(aes(x=no3.mgL, y=e.coli.cfu)) +
  geom_point(color = "black") +
   geom_smooth(method=lm, color = "black") +
  scale_x_continuous(trans = 'log10') +
  scale_y_continuous(trans = 'log10')

print(N_ECFU_no3)
```

#### Now run linear regression for N_ECFU_no3.
```{r}
N_ECFU_no3_lm <- lm(e.coli.cfu ~ no3.mgL, data=NORO)

summary(N_ECFU_no3_lm)
```

#### With *m = -213.1* and *p = 0.49*, there is a *negative*, *non-significant* correlation between *ECFU* and *no3*.

</br>

## BICO

</br>

### *Biological Indicator:*

#### Biological Vs Conductivity
```{r}
B_BS_Con <- BICO %>%
  ggplot(aes(x=conductivity.uscm, y=biological_score)) +
  geom_point(color = "red") +
   geom_smooth(method=lm, color = "red") +
  scale_x_continuous(trans = 'log10') +
  scale_y_continuous(trans = 'log10')

print(B_BS_Con)
```

#### Now run linear regression for B_BS_Con.
```{r}
B_BS_Con_lm <- lm(biological_score ~ conductivity.uscm, data=BICO)

summary(B_BS_Con_lm)
```

#### With *m = 0.0046* and *p = 0.86*, there is a *positive*, *non-significant* correlation between *biological score* and *conductivity*.

</br>

#### Biological Vs Turbidity
```{r}
B_BS_Turb <- BICO %>%
  ggplot(aes(x=turbidity.ntu, y=biological_score)) +
  geom_point(color = "blue") +
   geom_smooth(method=lm, color = "blue") +
  scale_x_continuous(trans = 'log10') +
  scale_y_continuous(trans = 'log10')

print(B_BS_Turb)
```

#### Now run linear regression for B_BS_Turb.
```{r}
B_BS_Turb_lm <- lm(biological_score ~ turbidity.ntu, data=BICO)

summary(B_BS_Turb_lm)
```

#### With *m = -0.13* and *p = 0.44*, there is a *negative*, *non-significant* correlation between *biological score* and *turbidity*.

</br>

#### Biological Vs Nitrate Concentration
```{r}
B_BS_no3 <- BICO %>%
  ggplot(aes(x=no3.mgL, y=biological_score)) +
  geom_point(color = "green") +
   geom_smooth(method=lm, color = "green") +
  scale_x_continuous(trans = 'log10') +
  scale_y_continuous(trans = 'log10')

print(B_BS_no3)
```

#### Now run linear regression for B_BS_no3.
```{r}
B_BS_no3_lm <- lm(biological_score ~ no3.mgL, data=BICO)

summary(B_BS_no3_lm)
```

#### With *m = -0.30* and *p = 0.84*, there is a *negative*, *non-significant* correlation between *biological score* and *no3*.

</br>

#### Biological Vs E. Coli CFU
```{r}
B_BS_ECFU <- BICO %>%
  ggplot(aes(x=e.coli.cfu, y=biological_score)) +
  geom_point(color = "purple") +
   geom_smooth(method=lm, color = "purple") +
  scale_x_continuous(trans = 'log10') +
  scale_y_continuous(trans = 'log10')

print(B_BS_ECFU)
```

#### Now run linear regression for B_BS_ECFU.
```{r}
B_BS_ECFU_lm <- lm(biological_score ~ e.coli.cfu, data=BICO)

summary(B_BS_ECFU_lm)
```

#### With *m = -0.00041* and *p = 0.19*, there is no relationship between *biological score* and *ECFU*.

</br>

### *E. Coli*
</br>
#### E. Coli Vs Conductivity
```{r}
B_ECFU_Con <- BICO %>%
  ggplot(aes(x=conductivity.uscm, y=e.coli.cfu)) +
  geom_point(color = "orange") +
   geom_smooth(method=lm, color = "orange") +
  scale_x_continuous(trans = 'log10') +
  scale_y_continuous(trans = 'log10')

print(B_ECFU_Con)
```

#### Now run linear regression for B_ECFU_Con.
```{r}
B_ECFU_Con_lm <- lm(e.coli.cfu ~ conductivity.uscm, data=BICO)

summary(B_ECFU_Con_lm)
```

#### With *m = 2.28* and *p = 0.90*, there is a *positive*, *non-significant* correlation between *ECFU* and *conductivity*.

</br>

#### E. Coli Vs Turbidity
```{r}
B_ECFU_Turb <- BICO %>%
  ggplot(aes(x=turbidity.ntu, y=e.coli.cfu)) +
  geom_point(color = "black") +
   geom_smooth(method=lm, color = "black") +
  scale_x_continuous(trans = 'log10') +
  scale_y_continuous(trans = 'log10')

print(B_ECFU_Turb)
```

#### Now run linear regression for B_ECFU_Turb.
```{r}
B_ECFU_Turb_lm <- lm(e.coli.cfu ~ turbidity.ntu, data=BICO)

summary(B_ECFU_Turb_lm)
```

#### With *m = 261.5* and *p = 0.01*, there is a *positive*, *significant* correlation between *ECFU* and *turbidity*.

</br>

#### E. Coli Vs Nitrate Concentration
```{r}
B_ECFU_no3 <- BICO %>%
  ggplot(aes(x=no3.mgL, y=e.coli.cfu)) +
  geom_point(color = "black") +
   geom_smooth(method=lm, color = "black") +
  scale_x_continuous(trans = 'log10') +
  scale_y_continuous(trans = 'log10')

print(B_ECFU_no3)
```

#### Now run linear regression for B_ECFU_no3.
```{r}
B_ECFU_no3_lm <- lm(e.coli.cfu ~ no3.mgL, data=BICO)

summary(B_ECFU_no3_lm)
```

#### With *m = -983.6* and *p = 0.33*, there is a *negative*, *non-significant* correlation between *ECFU* and *no3*.

</br>

### In summary, exploration of correlation of stream health indicators showed that most interactions were non-significant. However, it is to be noted that the ECFU and turbidity interaction was significantly positive. In further analysis, this may serve as a check to show that correlative analyses are robust enough with data provided. In aquatic microbiology, the positive relationship between ECFU growth and turbidity is generally accepted because the more bacteria that grow in the water column, the cloudier it will be.

### Additionally, there was a weird significant correlation between biological score and conductivity at the NORO site only. The relationship is negative, meaning that biological score decreased as conductivity increases, which make sense. However, NORO was the only site to show this interaction as significant.

### 2021-10-08 code/processing_code/processingscript.R

@@ -4,63 +4,133 @@
 #this script loads the raw data, processes and cleans it 
 #and saves it as Rds file in the processed_data folder
 
-#load needed packages. make sure they are installed.
-library(readxl) #for loading Excel files
-library(dplyr) #for data processing
-library(here) #to set paths
+#load needed packages. Install any that need to be installed.
+
+library(readxl) 
+library(dplyr) 
+library(here)
+library(tidyverse)
+library(tidyr)
+library(readr)
+library(stringr)
+library(naniar)
 
 #path to data
 #note the use of the here() package and not absolute paths
-data_location <- here::here("data","raw_data","exampledata.xlsx")
+data_location <- here::here("data","raw_data","UOWN_data_master_spring2021 (2).xlsx")
 
 #load data. 
-#note that for functions that come from specific packages (instead of base R)
-# I often specify both package and function like so
-#package::function() that's not required one could just call the function
-#specifying the package makes it clearer where the function "lives",
-#but it adds typing. You can do it either way.
 rawdata <- readxl::read_excel(data_location)
 
 #take a look at the data
 dplyr::glimpse(rawdata)
 
-#dataset is so small, we can print it to the screen.
-#that is often not possible.
+#or
+
 print(rawdata)
 
-# looks like we have measurements for height (in centimeters) and weight (in kilogram)
+#I had to do a little digging, but it looks like R has not registered values
+#designated "NA" as true missing values. We will have to fix that real quick.
 
-# there are some problems with the data: 
-# There is an entry which says "sixty" instead of a number. 
-# Does that mean it should be a numeric 60? It somehow doesn't make
-# sense since the weight is 60kg, which can't happen for a 60cm person (a baby)
-# Since we don't know how to fix this, we need to remove the person.
-# This "sixty" entry also turned all Height entries into characters instead of numeric.
-# We need to fix that too.
-# Then there is one person with a height of 6. 
-# that could be a typo, or someone mistakenly entered their height in feet.
-# Since we unfortunately don't know, we'll have to remove this person.
-# similarly, there is a person with weight of 7000, which is impossible,
-# and one person with missing weight.
-# to be able to analyze the data, we'll remove those 5 individuals
+rawdata <- rawdata %>%
+  replace_with_na(replace = list(WSID = "NA",	
[…]

### 2021-10-08 products/manuscript/Manuscript.Rmd
Project Part 2 - CC

---
title: "Relationship Between Common Stream Health Indicators"
subtitle: ""
author: Carter Coleman
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
documentclass: article
output:
  bookdown::word_document2: 
    toc: false
    number_sections: true
  bookdown::html_document2: 
    toc: false
bibliography: ../references.bib
csl: ../apa.csl
---

# Summary/Abstract

# Introduction

_When monitoring freshwater streams, stream health is typically measured by its physical characteristics (chemistry, surrounding land use area, channelization, etc.); Microbial (fecal indicator), and ecology (diversity of indicator macroinvertebrates). However, these categories are usually analyzed separately to assess stream health. As such, my question is how well are these parameters correlated with each other. Specifically, can you use one of the indicators mentioned (physical, microbial, and ecological) to describe both of the others to make a statement about overall stream health?_

_I would expect that because the all three are correlated with urbanization/anthropocentric impact, a negative impact in one indicator category will indicate a correlated negative impact in the other categories. The only caveat is that because level of noise in stream health data, significant correlations may be difficult to find._

# Methods and Results

_UOWN has provided a great deal of data dating back to 2003. First,I will need to wrangle missing data values that are a result of inconsistent sampling efforts. This will make the data easier to work with and narrow down a time frame for the data that is consistent and free of sampling bias. Next, I will analyze and compare trends in the change of chemical (conductivity, turbidity,	no3.mgL); ecological(Biological Score); and microbial indicator data (e.coli CFU). Generally speaking, increases in chemical properties like conductivity, turbidity,	no3.mgL and microbial indicator abundance (i.e. ecoli CFU) indicate decreased stream health. While a decrease in biological score indicates decrease in stream health. Next, since biological scores are used as a measure of broader ecological health, I will run t-tests on if the other factors significantly effect biological score. Finally, I will test the interaction of all three via multivariate analysis._

## Data aquisition
_The data I am using is the publically avaialable data provided by the Upper Oconee Watershed Network (UOWN). This data is part of a citizen science training effort for routine watershed-scale monitoring of the upper oconee river watershed. Trained freshwater scientists assist in training members of the community to collect both chemical and biological stream characteristics that are useful in describing stream health. All data is compliled and posted in raw form on the UOWN website. Of interest to me are the following variables: WSID (ID number),	biological_score,	conductivity, turbidity,	no3.mgL,	e.coli.cfu,	month,	year,	day. The raw data shows 3,075 observations before cleaning. However, while cleaning I had to eliminate "po4.mgl" and "temperature" due to significant amounts of missing data._

#### Load import packages:
```{r}
library(tidyverse)
library(readxl)
```

#### Lodaing data and providing a summary for the raw data
```{r, warning=FALSE}
data_location <- here::here("data/raw_data/UOWN_data_master_spring2021 (2).xlsx")

rawdata <- readxl::read_excel(data_location)

glimpse(rawdata)
```

## Data import and cleaning

## Exploratory analysis

## Full analysis

# Discussion

## Summary and Interpretation

## Strengths and Limitations

## Conclusions

# References

### 2021-10-08 code/analysis_code/analysisscript.R

@@ -1,69 +1,8 @@
 ###############################
 # analysis script
 #
-#this script loads the processed, cleaned data, does a simple analysis
-#and saves the results to the results folder
+#this script loads the processed, cleaned data, does complex analysis
+#for Carter Coleman's UOWN indicator of Stream Health Project.
 
-#load needed packages. make sure they are installed.
-library(ggplot2) #for plotting
-library(broom) #for cleaning up output from lm()
-library(here) #for data loading/saving
-
-#path to data
-#note the use of the here() package and not absolute paths
-data_location <- here::here("data","processed_data","processeddata.rds")
-
-#load data. 
-mydata <- readRDS(data_location)
-
-######################################
-#Data exploration/description
-######################################
-#I'm using basic R commands here.
-#Lots of good packages exist to do more.
-#For instance check out the tableone or skimr packages
-
-#summarize data 
-mysummary = summary(mydata)
-
-#look at summary
-print(mysummary)
-
-#do the same, but with a bit of trickery to get things into the 
-#shape of a data frame (for easier saving/showing in manuscript)
-summary_df = data.frame(do.call(cbind, lapply(mydata, summary)))
-
-#save data frame table to file for later use in manuscript
-summarytable_file = here("results", "summarytable.rds")
-saveRDS(summary_df, file = summarytable_file)
-
-
-#make a scatterplot of data
-#we also add a linear regression line to it
-p1 <- mydata %>% ggplot(aes(x=Height, y=Weight)) + geom_point() + geom_smooth(method='lm')
-
-#look at figure
-plot(p1)
-
-#save figure
-figure_file = here("results","resultfigure.png")
-ggsave(filename = figure_file, plot=p1) 
-
-######################################
-#Data fitting/statistical analysis
-######################################
-
-# fit linear model
-lmfit <- lm(Weight ~ Height, mydata)  
-
-# place results from fit into a data frame with the tidy function
-lmtable <- broom::tidy(lmfit)
-
-#look at fit results
-print(lmtable)
-
-# save fit results table  
-table_file = here("results", "resulttable.rds")
-saveRDS(lmtable, file = table_file)
 
   

### 2021-10-09 code/processing_code/Data_Exploration.Rmd
Project Part 2 - CC

@@
 ```
- <- source("processingscript.R", local = knitr::knit_global())
 

### 2021-10-14 code/processing_code/Data_Exploration.Rmd
Fixed Processing Script - CC

@@
 ```{r}
-MIDO <- readRDS("../../data/processed_data/MIDO.rds")
-NORO <- readRDS("../../data/processed_data/NORO.rds")
-BICO <- readRDS("../../data/processed_data/BICO.rds")
+MIDO <- here::here("data","processed_data","MIDO.RDS")
+NORO <- here::here("data","processed_data","NORO.RDS")
+BICO <- here::here("data","processed_data","BICO.RDS")
+
+readRDS(MIDO)
+readRDS(NORO)
+readRDS(BICO)
 ```

### 2021-10-14 code/processing_code/processingscript.R

@@ -14,6 +14,7 @@ library(tidyr)
 library(readr)
 library(stringr)
 library(naniar)
+library(here)
 
 #path to data
 #note the use of the here() package and not absolute paths
@@ -106,14 +107,11 @@ summary(data_cleaning)
 
 data_cleaning$stream_ID = substr(data_cleaning$WSID,1,4)
 
-MIDO <- data_cleaning %>%
-  filter( , stream_ID == "MIDO")
+MIDO <- filter(data_cleaning, stream_ID == "MIDO")
 
-NORO <- data_cleaning %>%
-  filter( , stream_ID == "NORO")
+NORO <- filter(data_cleaning, stream_ID == "NORO")
 
-BICO <- data_cleaning %>%
-  filter( , stream_ID == "BICO")
+BICO <- filter(data_cleaning, stream_ID == "BICO")
 
 summary(MIDO)
 summary(NORO)
@@ -129,8 +127,15 @@ summary(BICO)
 #See "analysis_code" folder for exploratory analysis.
 
 #Save files
-saveRDS(MIDO, "../../data/processed_data/MIDO.rds")
-saveRDS(NORO, "../../data/processed_data/NORO.rds")
-saveRDS(BICO, "../../data/processed_data/BICO.rds")
+MIDO_location <- here::here("data","processed_data","MIDO.rds")
+NORO_location <- here::here("data","processed_data","NORO.rds")
+BICO_location <- here::here("data","processed_data","BICO.rds")
+
+saveRDS(MIDO, file = MIDO_location)
+saveRDS(NORO, file = NORO_location)
+saveRDS(BICO, file = BICO_location)
+
+
+
 
 

### 2021-10-14 code/processing_code/Data_Exploration.Rmd
Fixed Issues in processing script - CC

--- 2021-10-09 code/processing_code/Data_Exploration.Rmd
@@
 ```{r}
-MIDO <- readRDS("../../data/processed_data/MIDO.rds")
-NORO <- readRDS("../../data/processed_data/NORO.rds")
-BICO <- readRDS("../../data/processed_data/BICO.rds")
+MIDO_location <- here::here("data","processed_data","MIDO.RDS")
+NORO_location <- here::here("data","processed_data","NORO.RDS")
+BICO_location <- here::here("data","processed_data","BICO.RDS")
+
+MIDO <- readRDS(MIDO_location)
+NORO <- readRDS(NORO_location)
+BICO <- readRDS(BICO_location)
 ```

### 2021-10-29 code/analysis_code/Project_Code_Dump.Rmd
UOWN_Analysis_Draft-CC

--- 2021-10-09 code/processing_code/Data_Exploration.Rmd
@@
 ---
-title: "Data Exploration"
+title: "Code_Dump_Project"
 author: "C. Coleman"
-date: "10/7/2021"
-output: word_document
+date: "10/28/2021"
+output: html_document
 ---
-
-# The following is an exploration of Stream Health Indicator correlation and comparison. The exploration will be structured as follows:
-### Each dataframe (MIDO, NORO, and BICO) will be analyzed separately.
-Within each dataframe, there are data explaining biological indicator values, stream chemistry, and fecal indicators over time.
-The only group with multiple components is stream chemistry.
-I plan to do a series of correlative analysis showing if there is some sort of relationship between the three indicator types to help inform further analysis via principle component, multivariate analysis, etc.
-Therfore, the goal of this exploration is basically to summarize how these indicators correlate.
-
+_Additionally, we will eventually be removing outliers from a series of relationships between many predictors and one outcome.
+You can imagine that because we are looking at many predictors with one outcome, we will either have some overlap in the outcome values we are trying to remove or potentially remove every single outcome value if we try to do this in the same data set.
+As such, we will need to split the MIDO, NORO, and BICO dataframes into individual dataframes that contain our outcome of interest and each single variable we are trying to assess a relationship.
+We will do this below._
+```{r, MIDO}
+bs_con_df <- MIDO %>%
+  select("biological_score", "conductivity.uscm")
+bs_ecfu_df <- MIDO %>%
+  select("biological_score", "e.coli.cfu")
+bs_no3_df <- MIDO %>%
+  select("biological_score", "no3.mgL")
+bs_turb_df <- MIDO %>%
+  select("biological_score", "turbidity.ntu")
+ecfu_con_df <- MIDO %>%
+  select("e.coli.cfu", "conductivity.uscm")
+ecfu_no3_df <- MIDO %>%
+  select("e.coli.cfu", "no3.mgL")
+ecfu_turb_df <- MIDO %>%
+  select("e.coli.cfu", "turbidity.ntu")
+```
 </br>
-
-#### First, we need to load all the libraries we will need.
-```{r}
-library(dplyr) 
-library(here)
-library(tidyverse)
-library(tidyr)
-library(readr)
-library(stringr)
+```{r, NORO}
+n_bs_con_df <- NORO %>%
+  select("biological_score", "conductivity.uscm")
+n_bs_ecfu_df <- NORO %>%
+  select("biological_score", "e.coli.cfu")
+n_bs_no3_df <- NORO %>%
+  select("biological_score", "no3.mgL")
+n_bs_turb_df <- NORO %>%
+  select("biological_score", "turbidity.ntu")
+n_ecfu_con_df <- NORO %>%
+  select("e.coli.cfu", "conductivity.uscm")
+n_ecfu_no3_df <- NORO %>%
+  select("e.coli.cfu", "no3.mgL")
+n_ecfu_turb_df <- NORO %>%
+  select("e.coli.cfu", "turbidity.ntu")
 ```
-
-#### Loading the data from the processing script. Note that because you are loading a library at the beginning, it may either take a while or take several tries.
-```{r}
-MIDO <- readRDS("../../data/processed_data/MIDO.rds")
-NORO <- readRDS("../../data/processed_data/NORO.rds")
-BICO <- readRDS("../../data/processed_data/BICO.rds")
-```
-
-#### take a look at the data frames to make sure everything worked correctly.
-```{r}
-view(MIDO)
-view(NORO)
-view(BICO)
+</br>
+```{r, BICO}
+b_bs_con_df <- BICO %>%
+  select("biological_score", "conductivity.uscm")
+b_bs_ecfu_df <- BICO %>%
+  select("biological_score", "e.coli.cfu")
+b_bs_no3_df <- BICO %>%
+  select("biological_score", "no3.mgL")
+b_bs_turb_df <- BICO %>%
+  select("biological_score", "turbidity.ntu")
+b_ecfu_con_df <- BICO %>%
+  select("e.coli.cfu", "conductivity.uscm")
+b_ecfu_no3_df <- BICO %>%
+  select("e.coli.cfu", "no3.mgL")
+b_ecfu_turb_df <- BICO %>%
+  select("e.coli.cfu", "turbidity.ntu")
 ```
@@
 
-## MIDO
+# Remove Outliers
+_To remove outliers, we must first create a new vector we will call "outliers_relationship", then we will just subtract it out of the original data frames (MIDO, NORO, BICO)_
+```{r, MIDO}
+Outliers_MIDO_bs_con <- as.numeric(names(bs_con_cd)[(bs_con_cd > (4/sample_size_MIDO))])
+Outliers_MIDO_bs_ecfu <- as.numeric(names(bs_ecfu_cd)[(bs_ecfu_cd > (4/sample_size_MIDO))])
+Outliers_MIDO_bs_no3 <- as.numeric(names(bs_no3_cd)[(bs_no3_cd > (4/sample_size_MIDO))])
+Outliers_MIDO_bs_turb <- as.numeric(names(bs_turb_cd)[(bs_turb_cd > (4/sample_size_MIDO))])
+Outliers_MIDO_ecfu_con <- as.numeric(names(ecfu_con_cd)[(ecfu_con_cd > (4/sample_size_MIDO))])
+Outliers_MIDO_ecfu_no3 <- as.numeric(names(ecfu_no3_cd)[(ecfu_no3_cd > (4/sample_size_MIDO))])
+Outliers_MIDO_ecfu_turb <- as.numeric(names(ecfu_turb_cd)[(ecfu_turb_cd > (4/sample_size_MIDO))])
+```
+</br>
+```{r, NORO}
+Outliers_NORO_bs_con <- as.numeric(names(n_bs_con_cd)[(n_bs_con_cd > (4/sample_size_NORO))])
+Outliers_NORO_bs_ecfu <- as.numeric(names(n_bs_ecfu_cd)[(n_bs_ecfu_cd > (4/sample_size_NORO))])
+Outliers_NORO_bs_no3 <- as.numeric(names(n_bs_no3_cd)[(n_bs_no3_cd > (4/sample_size_NORO))])
+Outliers_NORO_bs_turb <- as.numeric(names(n_bs_turb_cd)[(n_bs_turb_cd > (4/sample_size_NORO))])
+Outliers_NORO_ecfu_con <- as.numeric(names(n_ecfu_con_cd)[(n_ecfu_con_cd > (4/sample_size_NORO))])
+Outliers_NORO_ecfu_no3 <- as.numeric(names(n_ecfu_no3_cd)[(n_ecfu_no3_cd > (4/sample_size_NORO))])
+Outliers_NORO_ecfu_turb <- as.numeric(names(n_ecfu_turb_cd)[(n_ecfu_turb_cd > (4/sample_size_NORO))])
+```
+</br>
+```{r, BICO}
+Outliers_BICO_bs_con <- as.numeric(names(b_bs_con_cd)[(b_bs_con_cd > (4/sample_size_BICO))])
+Outliers_BICO_bs_ecfu <- as.numeric(names(b_bs_ecfu_cd)[(b_bs_ecfu_cd > (4/sample_size_BICO))])
+Outliers_BICO_bs_no3 <- as.numeric(names(b_bs_no3_cd)[(b_bs_no3_cd > (4/sample_size_BICO))])
+Outliers_BICO_bs_turb <- as.numeric(names(b_bs_turb_cd)[(b_bs_turb_cd > (4/sample_size_BICO))])
+Outliers_BICO_ecfu_con <- as.numeric(names(b_ecfu_con_cd)[(b_ecfu_con_cd > (4/sample_size_BICO))])
+Outliers_BICO_ecfu_no3 <- as.numeric(names(b_ecfu_no3_cd)[(b_ecfu_no3_cd > (4/sample_size_BICO))])
+Outliers_BICO_ecfu_turb <- as.numeric(names(b_ecfu_turb_cd)[(b_ecfu_turb_cd > (4/sample_size_BICO))])
+```
+</br>
+_With the outlier vectors created, all we need to do now is subtract them from the original cleaned data frames (MIDO, NORO, BICO)._
+```{r, MIDO}
+bs_con_df <- bs_con_df
+bs_ecfu_df <- bs_ecfu_df
+bs_no3_df <- bs_no3_df
+bs_turb_df <- bs_turb_df
+ecfu_con_df <- ecfu_con_df
+ecfu_no3_df <- ecfu_no3_df
+ecfu_turb_df <- ecfu_turb_df
+```
+</br>
+```{r, NORO}
+n_bs_con_df <- 
+n_bs_ecfu_df <- 
+n_bs_no3_df <- 
+n_bs_turb_df <- 
+n_ecfu_con_df <- 
+n_ecfu_no3_df <- 
+n_ecfu_turb_df <- 
+```
+</br>
+```{r, BICO}
+b_bs_con_df <- 
+b_bs_ecfu_df <- 
+b_bs_no3_df <- 
+b_bs_turb_df <- 
+b_ecfu_con_df <- 
+b_ecfu_no3_df <- 
+b_ecfu_turb_df <- 
+```
+```{r, MIDO}
+MIDO2 <- MIDO %>%
+  [-Outliers_MIDO_bs_con] 
+  MIDO[-Outliers_MIDO_bs_ecfu, ] 
+  MIDO[-Outliers_MIDO_bs_no3, ] 
+  MIDO[-Outliers_MIDO_bs_turb, ] 
+  MIDO[-Outliers_MIDO_ecfu_con, ] 
+  MIDO[-Outliers_MIDO_ecfu_no3, ] 
+  MIDO[-Outliers_MIDO_ecfu_turb, ]
+```
 
 </br>
-
-### *Biological Indicator:*
-
-#### Biological Vs Conductivity
+# Linear Regression Analysis Corrected for Outliers
+</br>
+## MIDO
+</br>
+#### Biological Indicator:
+_Biological Vs Conductivity_
 ```{r}
@@
 
+#### Now let's use the gridExtra package to view all the MIDO Biological Indicator outcome relationship graphs together.
+```{r}
+grid.arrange(BS_Con, BS_ECFU, BS_no3, BS_Turb, ncol = 2)
+```
+
 ### *E. Coli*
@@
 
+</br>
+
+#### Now let's use the gridExtra package to view all the MIDO E. coli CFU outcome relationship graphs together.
+```{r}
+grid.arrange(ECFU_Con, ECFU_no3, ECFU_Turb, ncol = 2)
+```
 </br>
@@
 
+</br>
+
+#### Now let's use the gridExtra package to view all the NORO Biological Indicator outcome relationship graphs together.
+```{r}
+grid.arrange(N_BS_Con, N_BS_ECFU, N_BS_no3, N_BS_Turb, ncol = 2)
+```
 </br>
@@
 #### With *m = -213.1* and *p = 0.49*, there is a *negative*, *non-significant* correlation between *ECFU* and *no3*.
+
+</br>
+
+#### Now let's use the gridExtra package to view all the NORO E. coli CFU outcome relationship graphs together.
+```{r}
+grid.arrange(N_ECFU_Con, N_ECFU_no3, N_ECFU_Turb, ncol = 2)
+```
 
@@
 
+#### Now let's use the gridExtra package to view all the BICO Biological Indicator outcome relationship graphs together.
+```{r}
+grid.arrange(B_BS_Con, B_BS_ECFU, B_BS_no3, B_BS_Turb, ncol = 2)
+```
+
+</br>
+
 ### *E. Coli*
@@
 
-### In summary, exploration of correlation of stream health indicators showed that most interactions were non-significant.
-However, it is to be noted that the ECFU and turbidity interaction was significantly positive.
-In further analysis, this may serve as a check to show that correlative analyses are robust enough with data provided.
-In aquatic microbiology, the positive relationship between ECFU growth and turbidity is generally accepted because the more bacteria that grow in the water column, the cloudier it will be.
-
-### Additionally, there was a weird significant correlation between biological score and conductivity at the NORO site only.
-The relationship is negative, meaning that biological score decreased as conductivity increases, which make sense.
-However, NORO was the only site to show this interaction as significant.
+#### Now let's use the gridExtra package to view all the BICO E. coli CFU outcome relationship graphs together.
+```{r}
+grid.arrange(B_ECFU_Con, B_ECFU_no3, B_ECFU_Turb, ncol = 2)
+```
+</br>

### 2021-10-29 code/analysis_code/analysis_uown.Rmd
UOWN_Analysis_Draft-CC

---
title: "analysis_uown"
author: "C. Coleman"
date: "10/27/2021"
output: html_document
---

# UOWN Data Analysis for Determining Correlation Between Stream Health Indicators.
</br>
_The following markdown conducts the statistical analysis of how different indicators of stream health can potentially predict each other. In the previous processing script, the main take away was that there were very few significant relationships between different chemical, fecal/microbial and biological indicators. The only significant relationships found between Turbity and E. Coli CFU. However, It struck me as peculiar to see biological indicator score and conductivity to have only one significant relationship. After further review of the exploratory figures, I now want to assess the data and relationships explored this far for impact of outliers, determine if it is appropriate to remove outlying values, and see if this helps find more significant relationships._

</br>

_Therefore, the following constitues the statistical analysis protocol that will be conducted in this markdown:_
</br>
#### PART 1: Outliers
-Assess each linear relationship previously performed for outliers in the data;
-Determine if it is appropriate to remove outliers;
-Rerun linear regressions for the different stream health indicator relationships and assess if more significant relationships exist.
</br>
#### Part 2: Linear Model Fit Analysis of MIDO Data

### Required packages
```{r}
library(tidyverse)
library(tidymodels)
library(here)
```

### Load Data
_Start by loading the cleaned data labeled as MIDO, NORO, and BICO into the Global Environment._
```{r}
MIDO_location <- here::here("data","processed_data","MIDO.RDS")
NORO_location <- here::here("data","processed_data","NORO.RDS")
BICO_location <- here::here("data","processed_data","BICO.RDS")

MIDO <- readRDS(MIDO_location)
NORO <- readRDS(NORO_location)
BICO <- readRDS(BICO_location)
```
</br>
_Additionally, I am going to start of by log fitting the E. coli data. This is because that data can have counts several orders of magnitude higher than the variables it is being assessed with. Therefore, model fit may be artificially bad.
```{r}
MIDO %>%
  mutate_at(vars(e.coli.cfu), ~log10(.))
NORO %>%
  mutate_at(vars(e.coli.cfu), ~log10(.))
BICO %>%
  mutate_at(vars(e.coli.cfu), ~log10(.))
```

</br>
### Load all the created linear models from the processing data markdown.
#### MIDO
_Biological Score_
```{r}
BS_Con_location <- here::here("data","processed_data","BS_Con.rds")
bs_con_lm <- readRDS(BS_Con_location)

BS_ECFU_location <- here::here("data","processed_data","BS_ECFU.rds")
bs_ecfu_lm <- readRDS(BS_ECFU_location)

BS_no3_location <- here::here("data","processed_data","BS_no3.rds")
bs_no3_lm <- readRDS(BS_no3_location)

BS_Turb_location <- here::here("data","processed_data","BS_Turb.rds")
bs_turb_lm <- readRDS(BS_Turb_location)
```
_ECFU_
```{r}
ECFU_Con_location <- here::here("data","processed_data","BS_Con.rds")
ecfu_con_lm <- readRDS(ECFU_Con_location)

ECFU_no3_location <- here::here("data","processed_data","BS_no3.rds")
ecfu_no3_lm <- readRDS(ECFU_no3_location)

ECFU_Turb_location <- here::here("data","processed_data","BS_Turb.rds")
ecfu_turb_lm <- readRDS(ECFU_Turb_location)
```

#### BICO
_Biological Score_
```{r}
B_BS_Con_location <- here::here("data","processed_data","B_BS_Con.rds")
b_bs_con_lm <- readRDS(B_BS_Con_location)

B_BS_ECFU_location <- here::here("data","processed_data","B_BS_ECFU.rds")
b_bs_ecfu_lm <- readRDS(B_BS_ECFU_location)

B_BS_no3_location <- here::here("data","processed_data","B_BS_no3.rds")
b_bs_no3_lm <- readRDS(B_BS_no3_location)

B_BS_Turb_location <- here::here("data","processed_data","B_BS_Turb.rds")
b_bs_turb_lm <- readRDS(B_BS_Turb_location)
```
_ECFU_
```{r}
B_ECFU_Con_location <- here::here("data","processed_data","B_BS_Con.rds")
b_ecfu_con_lm <- readRDS(B_ECFU_Con_location)

B_ECFU_no3_location <- here::here("data","processed_data","B_BS_no3.rds")
b_ecfu_no3_lm <- readRDS(B_ECFU_no3_location)

B_ECFU_Turb_location <- here::here("data","processed_data","B_BS_Turb.rds")
b_ecfu_turb_lm <- readRDS(B_ECFU_Turb_location)
```

#### NORO
_Biological Score_
```{r}
N_BS_Con_location <- here::here("data","processed_data","N_BS_Con.rds")
n_bs_con_lm <- readRDS(N_BS_Con_location)

N_BS_ECFU_location <- here::here("data","processed_data","N_BS_ECFU.rds")
n_bs_ecfu_lm <- readRDS(N_BS_ECFU_location)

N_BS_no3_location <- here::here("data","processed_data","N_BS_no3.rds")
n_bs_no3_lm <- readRDS(N_BS_no3_location)

N_BS_Turb_location <- here::here("data","processed_data","N_BS_Turb.rds")
n_bs_turb_lm <- readRDS(N_BS_Turb_location)
```
_ECFU_
```{r}
N_ECFU_Con_location <- here::here("data","processed_data","N_BS_Con.rds")
n_ecfu_con_lm <- readRDS(N_ECFU_Con_location)

N_ECFU_no3_location <- here::here("data","processed_data","N_BS_no3.rds")
n_ecfu_no3_lm <- readRDS(N_ECFU_no3_location)

N_ECFU_Turb_location <- here::here("data","processed_data","N_BS_Turb.rds")
n_ecfu_turb_lm <- readRDS(N_ECFU_Turb_location)
```

### Outlier identification using Box and Whisker Plots.
_Outliers in these data are going to be treated as errors in reporting. This is because the data is predominantly collected by non-scientist volunteers, who have had training with scientists. As such, it is likely that sampling of mis-identification of macroinvertebrates taxa has ocurred. Therefore, an argument can be made that assessing for anomalies in the data could help tease apart the true relationships that exist between different stream health indicators._
</br>
_As such, we will start by analyzing the data for outliers by evaluating relationships with cook's distance._
</br>
#### Set sample size for sample size parameter (4/n) for each data set (MIDO, NORO, BICO)
```{r}
sample_size_MIDO <- nrow(MIDO)
sample_size_NORO <- nrow(NORO)
sample_size_BICO <- nrow(BICO)
```
</br>
#### MIDO
</br>
_Outcome: Biological Score_
```{r}
bs_con_cd <- cooks.distance(bs_con_lm)
plot(bs_con_cd, pch="*", cex=2, main="MIDO: Conductivity vs. Biological Score with Outliers")
abline(h = 4/sample_size_MIDO, col="red")

bs_ecfu_cd <- cooks.distance(bs_ecfu_lm)
plot(bs_ecfu_cd, pch="*", cex=2, main="MIDO: E. coli CFU vs. Biological Score with Outliers")
abline(h = 4/sample_size_MIDO, col="red")

bs_no3_cd <- cooks.distance(bs_no3_lm)
plot(bs_no3_cd, pch="*", cex=2, main="MIDO: NO3 vs. Biological Score with Outliers")
abline(h = 4/sample_size_MIDO, col="red")

bs_turb_cd <- cooks.distance(bs_turb_lm)
plot(bs_turb_cd, pch="*", cex=2, main="MIDO: Turbidity vs. Biological Score with Outliers")
abline(h = 4/sample_size_MIDO, col="red")
```
</br>
_Outcome: E.coli CFU_
```{r}
ecfu_con_cd <- cooks.distance(ecfu_con_lm)
plot(ecfu_con_cd, pch="*", cex=2, main="MIDO: Conductivity vs. E. coli CFU with Outliers")
abline(h = 4/sample_size_MIDO, col="red")

ecfu_no3_cd <- cooks.distance(ecfu_no3_lm)
plot(ecfu_no3_cd, pch="*", cex=2, main="MIDO: NO3 vs. E. coli CFU with Outliers")
abline(h = 4/sample_size_MIDO, col="red")

ecfu_turb_cd <- cooks.distance(ecfu_turb_lm)
plot(ecfu_turb_cd, pch="*", cex=2, main="MIDO: Turbidity vs. E. coli CFU with Outliers")
abline(h = 4/sample_size_MIDO, col="red")
```
</br>
#### NORO
</br>
_Outcome: Biological Score_
```{r}
n_bs_con_cd <- cooks.distance(n_bs_con_lm)
plot(n_bs_con_cd, pch="*", cex=2, main="NORO: Conductivity vs. Biological Score with Outliers")
abline(h = 4/sample_size_NORO, col="red")

n_bs_ecfu_cd <- cooks.distance(n_bs_ecfu_lm)
plot(n_bs_ecfu_cd, pch="*", cex=2, main="NORO: E. coli CFU vs. Biological Score with Outliers")
abline(h = 4/sample_size_NORO, col="red")

n_bs_no3_cd <- cooks.distance(n_bs_no3_lm)
plot(n_bs_no3_cd, pch="*", cex=2, main="NORO: NO3 vs. Biological Score with Outliers")
abline(h = 4/sample_size_NORO, col="red")

n_bs_turb_cd <- cooks.distance(n_bs_turb_lm)
plot(n_bs_turb_cd, pch="*", cex=2, main="NORO: Turbidity vs. Biological Score with Outliers")
abline(h = 4/sample_size_NORO, col="red")
```
</br>
_Outcome: E.coli CFU_
```{r}
n_ecfu_con_cd <- cooks.distance(n_ecfu_con_lm)
plot(n_ecfu_con_cd, pch="*", cex=2, main="NORO: Conductivity vs. E. coli CFU with Outliers")
abline(h = 4/sample_size_NORO, col="red")

n_ecfu_no3_cd <- cooks.distance(n_ecfu_no3_lm)
plot(n_ecfu_no3_cd, pch="*", cex=2, main="NORO: NO3 vs. E. coli CFU with Outliers")
abline(h = 4/sample_size_NORO, col="red")

n_ecfu_turb_cd <- cooks.distance(n_ecfu_turb_lm)
plot(n_ecfu_turb_cd, pch="*", cex=2, main="NORO: Turbidity vs. E. coli CFU with Outliers")
abline(h = 4/sample_size_NORO, col="red")
```
</br>
#### BICO
</br>
_Outcome: Biological Score_
```{r}
b_bs_con_cd <- cooks.distance(b_bs_con_lm)
plot(b_bs_con_cd, pch="*", cex=2, main="BICO: Conductivity vs. Biological Score with Outliers")
abline(h = 4/sample_size_BICO, col="red")

b_bs_ecfu_cd <- cooks.distance(b_bs_ecfu_lm)
plot(b_bs_ecfu_cd, pch="*", cex=2, main="BICO: E. Coli CFU vs. Biological Score with Outliers")
abline(h = 4/sample_size_BICO, col="red")

b_bs_no3_cd <- cooks.distance(b_bs_no3_lm)
plot(b_bs_no3_cd, pch="*", cex=2, main="BICO: NO3 vs. Biological Score with Outliers")
abline(h = 4/sample_size_BICO, col="red")

b_bs_turb_cd <- cooks.distance(bs_turb_lm)
plot(b_bs_turb_cd, pch="*", cex=2, main="BICO: Turbidity vs. Biological Score with Outliers")
abline(h = 4/sample_size_BICO, col="red")
```
</br>
_Outcome: E.coli CFU_
```{r}
b_ecfu_con_cd <- cooks.distance(b_ecfu_con_lm)
plot(b_ecfu_con_cd, pch="*", cex=2, main="BICO: Conductivity vs. E. coli CFU with Outliers")
abline(h = 4/sample_size_BICO, col="red")

b_ecfu_no3_cd <- cooks.distance(b_ecfu_no3_lm)
plot(b_ecfu_no3_cd, pch="*", cex=2, main="BICO: NO3 vs. E. coli CFU with Outliers")
abline(h = 4/sample_size_BICO, col="red")

b_ecfu_turb_cd <- cooks.distance(b_ecfu_turb_lm)
plot(b_ecfu_turb_cd, pch="*", cex=2, main="BICO: Turbidity vs. E. coli CFU with Outliers")
abline(h = 4/sample_size_BICO, col="red")
```
</br>
_The results of the Cook's Distance show that while there are outliers, very few of them actually exceed a Cook's Distance value (CDV) of 1. The CDV value of 1 is significant because it is the statistical community consensus value that indicates an influential outlier. As such, it is my educated opinion to leave all data points in, as I do not believe they are skewing liner modle results in a significant enough way to warrant removal._
</br>

# Part 2: Linear Model Fit Assessment for MIDO Data
</br>
_At this point,the main take-aways from data analysis thus far is that correlation of stream health indicators showed that most interactions were non-significant, with the exception of E. coli (cfu) and turbidity. Additionally, the data set is most likely not skewed by influential outliers._
</br>
_Moving forward, I want to focus in on just the MIDO dataset. As you can tell, working with many variables in three different data sets can be quite cumbersome to read through. For ease of interpretation, the MIDO data set will serve as the data frame where I test how robust the linear fit models actually are for the data. MIDO is the only data set large enough that when we split the data set into a "test" and "train" data set, I actually have confidence in the resulting model fit analysis. For comparison, MIDO has an n = 88; NORO has an n = 35; and BICO has an n = 23. The only thing we are missing by dropping the other sites is assessing site-by-site comparison, which I am choosing not to focus on anyway._
</br>
#### Splitting MIDO Data
_As discussed above, we need to split the MIDO data into a test and train data set. We do this to see if we get approximately the same linear relationship between the larger (75% of MIDO) train data and the smaller (25% of MIDO) test data._
```{r}
data_split <- initial_split(MIDO, prop = 3/4)

# Create data frames for the two sets:
train_data <- training(data_split)
test_data  <- testing(data_split)
```
</br>
_Create a recipe for each linear model relationship in the MIDO dataframe._
```{r}
Recipe_bs_con <- recipe(biological_score ~ conductivity.uscm, data = MIDO)
Recipe_bs_ecfu <- recipe(biological_score ~ e.coli.cfu, data = MIDO)
Recipe_bs_no3 <- recipe(biological_score ~ no3.mgL, data = MIDO)
Recipe_bs_turb <- recipe(biological_score ~ turbidity.ntu, data = MIDO)
Recipe_ecfu_con <- recipe(e.coli.cfu ~ e.coli.cfu, data = MIDO)
Recipe_ecfu_no3 <- recipe(e.coli.cfu ~ no3.mgL, data = MIDO)
Recipe_ecfu_turb <- recipe(e.coli.cfu ~ turbidity.ntu, data = MIDO)
```
_With our seven linear model recipes created, we will define of linear regression model pipe._
```{r}
linear_mod <- linear_reg() %>% 
  set_engine("lm") %>%
  set_mode("regression")
```
_Next, create a workflow that adds our seven recipes and model together._
```{r}
#Biological Score and Conductivity
bs_con_wflow <- 
  workflow() %>% 
  add_model(linear_mod) %>% 
  add_recipe(Recipe_bs_con)

#Biological Score and E. coli(cfu)
bs_ecfu_wflow <- 
  workflow() %>% 
  add_model(linear_mod) %>% 
  add_recipe(Recipe_bs_ecfu)

#Biological Score and NO3
bs_no3_wflow <- 
  workflow() %>% 
  add_model(linear_mod) %>% 
  add_recipe(Recipe_bs_no3)

#Biological Score and Turbidity
bs_turb_wflow <- 
  workflow() %>% 
  add_model(linear_mod) %>% 
  add_recipe(Recipe_bs_turb)

#E. Coli (cfu) and Conductivity
ecfu_con_wflow <- 
  workflow() %>% 
  add_model(linear_mod) %>% 
  add_recipe(Recipe_ecfu_con)

#E. Coli (cfu) and NO3
ECFU_no3_wflow <- 
  workflow() %>% 
  add_model(linear_mod) %>% 
  add_recipe(Recipe_ecfu_no3)

#E. Coli (cfu) and Turbidity
ecfu_turb_wflow <- 
  workflow() %>% 
  add_model(linear_mod) %>% 
  add_recipe(Recipe_ecfu_turb)
```

## Modeling Train Data Predictions Using Workflows
_Using the workflow above, lets now fit each pf the seven stream health indicator linear models to the train data set._
</br>
_Defining a command that runs model fitting to each of the seven models_
```{r}
bs_con_fit <- 
  bs_con_wflow %>% 
  fit(data = train_data)

bs_ecfu_fit <- 
  bs_ecfu_wflow %>% 
  fit(data = train_data)

bs_no3_fit <- 
  bs_no3_wflow %>% 
  fit(data = train_data)

bs_turb_fit <- 
  bs_turb_wflow %>% 
  fit(data = train_data)

ecfu_con_fit <- 
  ecfu_con_wflow %>% 
  fit(data = train_data)

ecfu_no3_fit <- 
  ECFU_no3_wflow %>% 
  fit(data = train_data)

ecfu_turb_fit <- 
  ecfu_turb_wflow %>% 
  fit(data = train_data)
```
</br>
_Pull linear regression fit models using parsnip()._
```{r}
bs_con_fit %>% 
  extract_fit_parsnip() %>% 
  tidy()

bs_ecfu_fit %>% 
  extract_fit_parsnip() %>% 
  tidy()

bs_no3_fit %>% 
  extract_fit_parsnip() %>% 
  tidy()

bs_turb_fit %>% 
  extract_fit_parsnip() %>% 
  tidy()

ecfu_con_fit %>% 
  extract_fit_parsnip() %>% 
  tidy()

ecfu_no3_fit %>% 
  extract_fit_parsnip() %>% 
  tidy()

ecfu_turb_fit %>% 
  extract_fit_parsnip() %>% 
  tidy()
```
#####Analysis here#####
</br>
#### Using Workflows to Make Predictions in the train data set.
_This is important because we are basically going to try and recreate the actual MIDO data set using the test and trained data and our linear models. We will eventually compare the test data predictions with the train data predictions to assess if our model is actually a good estimator of the relationships between stream health indicators. Additionally, this prediction step allows us to "artificially" create a larger n for our data, allowing fit to be assessed more accurately. This will be done using the augment() function to make predicted outcomes in the test and train data._
```{r}
bs_con_aug <- 
  augment(bs_con_fit, train_data)

bs_ecfu_aug <- 
  augment(bs_ecfu_fit, train_data)

bs_no3_aug <- 
  augment(bs_no3_fit, train_data)

bs_turb_aug <- 
  augment(bs_turb_fit, train_data)

ecfu_con_aug <- 
  augment(ecfu_con_fit, train_data)

ecfu_no3_aug <- 
  augment(ecfu_no3_fit, train_data)

ecfu_turb_aug <- 
  augment(ecfu_turb_fit, train_data)
```  
</br>
#### Using Root Mean Squared Error (RMSE) to Assess Model Fit.

_Assessing fit for train data predictions by calculating RMSE for the test data predictions of stream health indicator linear relationships._
```{r}
bs_con_aug %>%
  rmse(truth = biological_score, .pred)

bs_ecfu_aug %>%
  rmse(truth = biological_score, .pred)

bs_no3_aug %>%
  rmse(truth = biological_score, .pred)

bs_turb_aug %>%
  rmse(truth = biological_score, .pred)

ecfu_con_aug %>%
  rmse(truth = e.coli.cfu, .pred)

ecfu_no3_aug %>%
  rmse(truth = e.coli.cfu, .pred)

ecfu_turb_aug %>%
  rmse(truth = e.coli.cfu, .pred)
```
</br>
#####Analysis here#####
</br>
## Modeling Test Data Predictions Using Workflows
_Using the workflow above, lets now fit each pf the seven stream health indicator linear models to the test data set._
</br>
_Defining a command that runs model fitting to each of the seven models_
```{r}
bs_con_fit_test <- 
  bs_con_wflow %>% 
  fit(data = test_data)

bs_ecfu_fit_test <- 
  bs_ecfu_wflow %>% 
  fit(data = test_data)

bs_no3_fit_test <- 
  bs_no3_wflow %>% 
  fit(data = test_data)

bs_turb_fit_test <- 
  bs_turb_wflow %>% 
  fit(data = test_data)

ecfu_con_fit_test <- 
  ecfu_con_wflow %>% 
  fit(data = test_data)

ecfu_no3_fit_test <- 
  ECFU_no3_wflow %>% 
  fit(data = test_data)

ecfu_turb_fit_test <- 
  ecfu_turb_wflow %>% 
  fit(data = test_data)
```
</br>
_Pull linear regression fit models using parsnip()._
```{r}
bs_con_fit_test %>% 
  extract_fit_parsnip() %>% 
  tidy()

bs_ecfu_fit_test %>% 
  extract_fit_parsnip() %>% 
  tidy()

bs_no3_fit_test %>% 
  extract_fit_parsnip() %>% 
  tidy()

bs_turb_fit_test %>% 
  extract_fit_parsnip() %>% 
  tidy()

ecfu_con_fit_test %>% 
  extract_fit_parsnip() %>% 
  tidy()

ecfu_no3_fit_test %>% 
  extract_fit_parsnip() %>% 
  tidy()

ecfu_turb_fit_test %>% 
  extract_fit_parsnip() %>% 
  tidy()
```
#####Analysis here#####
</br>
#### Using Workflows to Make Predictions in the test data set.
_This is important because we are basically going to try and recreate the actual MIDO data set using the test and trained data and our linear models. This prediction step allows us to "artificially" create a larger n for our data, allowing fit to be assessed more accurately for data predicted from split data. This will be done using the augment() function to make predicted outcomes in the test and train data._
```{r}
bs_con_aug_test <- 
  augment(bs_con_fit, test_data)

bs_ecfu_aug_test <- 
  augment(bs_ecfu_fit, test_data)

bs_no3_aug_test <- 
  augment(bs_no3_fit, test_data)

bs_turb_aug_test <- 
  augment(bs_turb_fit, test_data)

ecfu_con_aug_test <- 
  augment(ecfu_con_fit, test_data)

ecfu_no3_aug_test <- 
  augment(ecfu_no3_fit, test_data)

ecfu_turb_aug_test <- 
  augment(ecfu_turb_fit, test_data)
```  
</br>
#### Using Root Mean Squared Error (RMSE) to Assess Model Fit.

_Assessing fit for train data predictions by calculating RMSE for the test data predictions of stream health indicator linear relationships._
```{r}
bs_con_aug_test %>%
  rmse(truth = biological_score, .pred)

bs_ecfu_aug_test %>%
  rmse(truth = biological_score, .pred)

bs_no3_aug_test %>%
  rmse(truth = biological_score, .pred)

bs_turb_aug_test %>%
  rmse(truth = biological_score, .pred)

ecfu_con_aug_test %>%
  rmse(truth = e.coli.cfu, .pred)

ecfu_no3_aug_test %>%
  rmse(truth = e.coli.cfu, .pred)

ecfu_turb_aug_test %>%
  rmse(truth = e.coli.cfu, .pred)
```
#####Analysis Here#####

### 2021-10-29 code/analysis_code/analysis_uown.Rmd
Project Part 3 - CC

@@
 However, It struck me as peculiar to see biological indicator score and conductivity to have only one significant relationship.
-After further review of the exploratory figures, I now want to assess the data and relationships explored this far for impact of outliers, determine if it is appropriate to remove outlying values, and see if this helps find more significant relationships._
+After further review of the exploratory figures, I now want to assess the data and relationships explored thus far for impact of outliers, determine if it is appropriate to remove outlying values, and see if this helps find more significant relationships._
+_The second part will include test/train splitting to predict stream health indicator values in the MIDO data set and assess the fit of linear models to the MIDO data.
+This is a more depth recreation of the linear model exploration.
+Obviously, we found that there weren't any significant relationships in exploratory analysis.
+However, this gives us a chance to assess how good a fit the models are in order to see if its even worth looking at the relationships._
 
@@
 
-_Therefore, the following constitues the statistical analysis protocol that will be conducted in this markdown:_
+_Therefore, the following constitutes the statistical analysis protocol that will be conducted in this markdown:_
 </br>
-#### PART 1: Outliers
--Assess each linear relationship previously performed for outliers in the data;
--Determine if it is appropriate to remove outliers;
--Rerun linear regressions for the different stream health indicator relationships and assess if more significant relationships exist.
+#### PART 1: Outliers in stream health indicator (Biological Score, E. Coli (cfu), Conductivity, NO3 (mg/L), and Turbidity) linear relationships.
+_1. Assess each linear relationship previously performed for outliers in the data;_
+_2. Determine if it is appropriate to remove outliers;_
+_3. Rerun linear regressions for the different stream health indicator relationships and assess if more significant relationships exist._
 </br>
 #### Part 2: Linear Model Fit Analysis of MIDO Data
+_1. Conduct test/train data splitting for MIDO Data._
+_2.
+Conduct linear modeling for relationships between stream health indicator predictors (Biological Score, E.
+Coli (cfu), Conductivity, NO3 (mg/L), and Turbidity) in the form of a workflow for the Test and Train MIDO data._
+_3. Predict values for Test and Train MIDO data._
+_4.
+Assess how well the MIDO linear model suite of relationships (Biological Score, E.
+Coli (cfu), Conductivity, NO3 (mg/L), and Turbidity) fir the test and train data using Root Mean Square Error (rmse).
+The result will measure how good of a fit the created models are to the data._
 
@@
 ```{r}
+#Set the location of the desired dataframes
 MIDO_location <- here::here("data","processed_data","MIDO.RDS")
@@
 
+#Load in the Dataframes
 MIDO <- readRDS(MIDO_location)
@@
 ```{r}
+#Change the E.Coli (cfu) variable to log10 scale using mutate()
 MIDO %>%
@@
 ### Load all the created linear models from the processing data markdown.
+_We need to load the exploratory linear models created previously in order to assess them for outliers that impact the linear regressions.
+This is because the cooksdistance() function calls for a linear model input.
 #### MIDO
@@
 ```{r}
+#Loading exploratory linear models for MIDO data with Biological Score as Outcome variable.
 BS_Con_location <- here::here("data","processed_data","BS_Con.rds")
@@
 ```{r}
+#Loading exploratory linear models for MIDO data with E. coli (cfu) as Outcome variable.
 ECFU_Con_location <- here::here("data","processed_data","BS_Con.rds")
@@
 ```{r}
+#Loading exploratory linear models for BICO data with Biological Score as Outcome variable.
 B_BS_Con_location <- here::here("data","processed_data","B_BS_Con.rds")
@@
 ```{r}
+#Loading exploratory linear models for BICO data with E. coli (cfu) as Outcome variable.
 B_ECFU_Con_location <- here::here("data","processed_data","B_BS_Con.rds")
@@
 ```{r}
+#Loading exploratory linear models for NORO data with Biological Score as Outcome variable.
 N_BS_Con_location <- here::here("data","processed_data","N_BS_Con.rds")
@@
 ```{r}
+#Loading exploratory linear models for NORO data with E. coli (cfu) as Outcome variable.
 N_ECFU_Con_location <- here::here("data","processed_data","N_BS_Con.rds")
@@
 
-### Outlier identification using Box and Whisker Plots.
-_Outliers in these data are going to be treated as errors in reporting.
+### Outlier identification using Cook's Distance.
+_Outliers in these data could be potentially ttreated as errors in reporting.
 This is because the data is predominantly collected by non-scientist volunteers, who have had training with scientists.
-As such, it is likely that sampling of mis-identification of macroinvertebrates taxa has ocurred.
+As such, it is likely that sampling of mis-identification of macroinvertebrates taxa may occurred.
 Therefore, an argument can be made that assessing for anomalies in the data could help tease apart the true relationships that exist between different stream health indicators._
@@
 ```{r}
+#This makes an argument that makes the sum number of rows in a particular data set equal to the sample size of the data set and names it accordingly.
 sample_size_MIDO <- nrow(MIDO)
@@
 ```{r}
+#Creates a dataframe that stores the cook's distance analysis of influential data points.
 bs_con_cd <- cooks.distance(bs_con_lm)
@@
 ```{r}
+#Creates a dataframe that stores the cook's distance analysis of influential data points.
 ecfu_con_cd <- cooks.distance(ecfu_con_lm)
@@
 ```{r}
+#Creates a dataframe that stores the cook's distance analysis of influential data points.
 n_bs_con_cd <- cooks.distance(n_bs_con_lm)
@@
 ```{r}
+#Creates a dataframe that stores the cook's distance analysis of influential data points.
 n_ecfu_con_cd <- cooks.distance(n_ecfu_con_lm)
@@
 ```{r}
+#Creates a dataframe that stores the cook's distance analysis of influential data points.
 b_bs_con_cd <- cooks.distance(b_bs_con_lm)
@@
 ```{r}
+#Creates a dataframe that stores the cook's distance analysis of influential data points.
 b_ecfu_con_cd <- cooks.distance(b_ecfu_con_lm)
@@
 ```{r}
+#Split the data with 1:4 ratio.
 data_split <- initial_split(MIDO, prop = 3/4)
@@
 ```
-#####Analysis here#####
 </br>
@@
 ```
-#####Analysis Here#####
+#### Save significant Test and Train linear models to the results folder.
+```{r}
+#Create a location and save linear model for E. coli (cfu) in the MIDO train data.
+MIDO_ECFU_Con_fit_location <- here::here("results","MIDO_ECFU_Con_fit.rds")
+saveRDS(ecfu_con_fit, file = MIDO_ECFU_Con_fit_location)
+
+#Create a location and save linear model for E. coli (cfu) in the MIDO test data.
+MIDO_ECFU_Con_fit_test_location <- here::here("results","MIDO_ECFU_Con_fit_test.rds")
+saveRDS(ecfu_con_fit_test, file = MIDO_ECFU_Con_fit_test_location)
+```
+</br>
+#### Save significant Test and Train RMSE to the results folder.
+```{r}
+#Create a location and save RMSE for E. coli (cfu) in the MIDO train data.
+MIDO_ECFU_Con_RMSE_location <- here::here("results","MIDO_ECFU_Con_RMSE.rds")
+saveRDS(ecfu_con_aug, file = MIDO_ECFU_Con_RMSE_location)
+
+#Create a location and save linear RMSE for E. coli (cfu) in the MIDO test data.
+MIDO_ECFU_Con_RMSE_test_location <- here::here("results","MIDO_ECFU_Con_RMSE_test.rds")
+saveRDS(ecfu_con_aug_test, file = MIDO_ECFU_Con_RMSE_test_location)
+```

### 2021-10-29 code/processing_code/Data_Exploration.Rmd
Project Part 3 - CC

@@
 library(stringr)
+library(gridExtra)
 ```
@@
 ## MIDO
+_This section of regression analysis and plotting deals exclusively with the MIDO data set._
 
@@
 
-#### Biological Vs Conductivity
+#### Biological Vs Conductivity:
+_This plot shows the relationship between biological score as a function of conductivity._
 ```{r}
@@
 
-#### Biological Vs Turbidity
+#### Biological Vs Turbidity:
+_This plot shows the relationship between biological score as a function of turbidity._
 ```{r}
@@
 #### Biological Vs Nitrate Concentration
+_This plot shows the relationship between biological score as a function of nitrate (no3) concentration._
 ```{r}
@@
 #### Biological Vs E. Coli CFU
+_This plot shows the relationship between biological score as a function of E. Coli (cfu)._
 ```{r}
@@
 
+#### Now let's use the gridExtra package to view all the MIDO Biological Indicator outcome relationship graphs together.
+```{r}
+#Creates a plot with all the MIDO biological score relationships in it.
+MIDO_bs_lm <- grid.arrange(BS_Con, BS_ECFU, BS_no3, BS_Turb, ncol = 2)
+```
+</br>
+_Save MIDO Biological Score figures_
+```{r}
+#Stores the location I want to save the graphs to.
+MIDO_bs_lm_location <- here::here("results","MIDO_bs_lm.png")
+#Saves graphs to stored location in ,my working directory.
+ggsave(MIDO_bs_lm, file = MIDO_bs_lm_location)
+```
+</br>
 ### *E. Coli*
@@
 #### E. Coli Vs Conductivity
+_This plot shows the relationship between E. coli (cfu) as a function of Conductivity._
 ```{r}
@@
 #### E. Coli Vs Turbidity
+_This plot shows the relationship between E. coli as a function of Turbidity._
 ```{r}
@@
 #### E. Coli Vs Nitrate Concentration
+_This plot shows the relationship between E. coli (cfu) as a function of nitrate (no3) concentration._
 ```{r}
@@
 
+#### Now let's use the gridExtra package to view all the MIDO E. coli CFU outcome relationship graphs together.
+```{r}
+MIDO_ECFU_lm <- grid.arrange(ECFU_Con, ECFU_no3, ECFU_Turb, ncol = 2)
+```
+</br>
+_Save MIDO ECFU figures_
+```{r}
+#Stores the location I want to save the graphs to.
+MIDO_ECFU_lm_location <- here::here("results","MIDO_ECFU_lm.png")
+#Saves graphs to stored location in ,my working directory.
+ggsave(MIDO_ECFU_lm, file = MIDO_ECFU_lm_location)
+```
+</br>
 ## NORO
+_the following linear analysis deals exclusively with the NORO dataset._
 
@@
 #### Biological Vs Conductivity
+_This plot shows the relationship between Biological Score as a function of Conductivity._
 ```{r}
@@
 #### Biological Vs Turbidity
+_This plot shows the relationship between Biological Score as a function of Turbidity._
 ```{r}
@@
 #### Biological Vs Nitrate Concentration
+_This plot shows the relationship between Biological Score as a function of nitrate (no3) Concentration._
 ```{r}
@@
 
-#### Biological Vs E. Coli CFU
+#### Biological Vs E. Coli CFU\
+_This plot shows the relationship between Biological Score as a function of E. coli (cfu)._
 ```{r}
@@
 </br>
+#### Now let's use the gridExtra package to view all the NORO Biological Indicator outcome relationship graphs together.
+```{r}
+NORO_bs_lm <- grid.arrange(N_BS_Con, N_BS_ECFU, N_BS_no3, N_BS_Turb, ncol = 2)
+```
+</br>
+_Save NORO Biological Score figures_
+```{r}
+#Stores the location I want to save the graphs to.
+NORO_bs_lm_location <- here::here("results","NORO_bs_lm.png")
+#Saves graphs to stored location in ,my working directory.
+ggsave(NORO_bs_lm, file = NORO_bs_lm_location)
+```
+</br>
+</br>
 
@@
 #### E. Coli Vs Conductivity
+_This plot shows the relationship between E. Coli (cfu) as a function of Conductivity._
 ```{r}
@@
 #### E. Coli Vs Turbidity
+_This plot shows the relationship between E. Coli (cfu) as a function of Turbidity._
 ```{r}
@@
 #### E. Coli Vs Nitrate Concentration
+_This plot shows the relationship between E. coli as a function of nitrate (no3) concentration._
 ```{r}
@@
 
-## BICO
+#### Now let's use the gridExtra package to view all the NORO E. coli CFU outcome relationship graphs together.
+```{r}
+NORO_ECFU_lm <- grid.arrange(N_ECFU_Con, N_ECFU_no3, N_ECFU_Turb, ncol = 2)
+```
+</br>
+_Save NORO ECFU figures_
+```{r}
+#Stores the location I want to save the graphs to.
+NORO_ECFU_lm_location <- here::here("results","NORO_ECFU_lm.png")
+#Saves graphs to stored location in ,my working directory.
+ggsave(NORO_ECFU_lm, file = NORO_ECFU_lm_location)
+```
+
+</br>
+
+## BICO:
+_The following Linear Regression analysis deals exclusively with the BICO data set._
 
@@
 #### Biological Vs Conductivity
+_This plot shows the relationship between biological score as a function of conductivity._
 ```{r}
@@
 #### Biological Vs Turbidity
+_This plot shows the relationship between biological score as a function of turbidity._
 ```{r}
@@
 #### Biological Vs E. Coli CFU
+_This plot shows the relationship between biological score as a function of E. coli (cfu)._
 ```{r}
@@
 
+#### Now let's use the gridExtra package to view all the BICO Biological Indicator outcome relationship graphs together.
+```{r}
+BICO_bs_lm <- grid.arrange(B_BS_Con, B_BS_ECFU, B_BS_no3, B_BS_Turb, ncol = 2)
+```
+</br>
+_Save BICO biological score figures_
+```{r}
+#Stores the location I want to save the graphs to.
+BICO_bs_lm_location <- here::here("results","BICO_bs_lm.png")
+#Saves graphs to stored location in ,my working directory.
+ggsave(BICO_bs_lm, file = BICO_bs_lm_location)
+```
+
+</br>
+
 ### *E. Coli*
@@
 #### E. Coli Vs Conductivity
+_This plot shows the relationship between E. coli (cfu) as a function of conductivity._
 ```{r}
@@
 #### E. Coli Vs Turbidity
+_This plot shows the relationship between E. coli (cfu) as a function of turbidity._
 ```{r}
@@
 
+#### Now let's use the gridExtra package to view all the BICO E. coli CFU outcome relationship graphs together.
+```{r}
+BICO_ECFU_lm <- grid.arrange(B_ECFU_Con, B_ECFU_no3, B_ECFU_Turb, ncol = 2)
+```
+</br>
+_Save BICO ECFU figures_
+```{r}
+#Stores the location I want to save the graphs to.
+BICO_ECFU_lm_location <- here::here("results","BICO_ECFU_lm.png")
+#Saves graphs to stored location in ,my working directory.
+ggsave(BICO_ECFU_lm, file = BICO_ECFU_lm_location)
+```
+</br>
+### Save the exploratory linear models
+#### MIDO
+_Biological Score_
+```{r}
+BS_Con_location <- here::here("data","processed_data","BS_Con.rds")
+saveRDS(BS_Con_lm, file = BS_Con_location)
+
+BS_ECFU_location <- here::here("data","processed_data","BS_ECFU.rds")
+saveRDS(BS_ECFU_lm, file = BS_ECFU_location)
+
+BS_no3_location <- here::here("data","processed_data","BS_no3.rds")
+saveRDS(BS_no3_lm, file = BS_no3_location)
+
+BS_Turb_location <- here::here("data","processed_data","BS_Turb.rds")
+saveRDS(BS_Turb_lm, file = BS_Turb_location)
+```
+_ECFU_
+```{r}
+ECFU_Con_location <- here::here("data","processed_data","BS_Con.rds")
+saveRDS(ECFU_Con_lm, file = ECFU_Con_location)
+
+ECFU_no3_location <- here::here("data","processed_data","BS_no3.rds")
+saveRDS(ECFU_no3_lm, file = ECFU_no3_location)
+
+ECFU_Turb_location <- here::here("data","processed_data","BS_Turb.rds")
+saveRDS(ECFU_Turb_lm, file = ECFU_Turb_location)
+```
+
+#### BICO
+_Biological Score_
+```{r}
+B_BS_Con_location <- here::here("data","processed_data","B_BS_Con.rds")
+saveRDS(B_BS_Con_lm, file = B_BS_Con_location)
+
+B_BS_ECFU_location <- here::here("data","processed_data","B_BS_ECFU.rds")
+saveRDS(B_BS_ECFU_lm, file = B_BS_ECFU_location)
+
+B_BS_no3_location <- here::here("data","processed_data","B_BS_no3.rds")
+saveRDS(B_BS_no3_lm, file = B_BS_no3_location)
+
+B_BS_Turb_location <- here::here("data","processed_data","B_BS_Turb.rds")
+saveRDS(B_BS_Turb_lm, file = B_BS_Turb_location)
+```
+_ECFU_
+```{r}
+B_ECFU_Con_location <- here::here("data","processed_data","B_BS_Con.rds")
+saveRDS(B_ECFU_Con_lm, file = B_ECFU_Con_location)
+
+B_ECFU_no3_location <- here::here("data","processed_data","B_BS_no3.rds")
+saveRDS(B_ECFU_no3_lm, file = B_ECFU_no3_location)
+
+B_ECFU_Turb_location <- here::here("data","processed_data","B_BS_Turb.rds")
+saveRDS(B_ECFU_Turb_lm, file = B_ECFU_Turb_location)
+```
+
+#### NORO
+_Biological Score_
+```{r}
+N_BS_Con_location <- here::here("data","processed_data","N_BS_Con.rds")
+saveRDS(N_BS_Con_lm, file = N_BS_Con_location)
+
+N_BS_ECFU_location <- here::here("data","processed_data","N_BS_ECFU.rds")
+saveRDS(N_BS_ECFU_lm, file = N_BS_ECFU_location)
+
+N_BS_no3_location <- here::here("data","processed_data","N_BS_no3.rds")
+saveRDS(N_BS_no3_lm, file = N_BS_no3_location)
+
+N_BS_Turb_location <- here::here("data","processed_data","N_BS_Turb.rds")
+saveRDS(N_BS_Turb_lm, file = N_BS_Turb_location)
+```
+_ECFU_
+```{r}
+N_ECFU_Con_location <- here::here("data","processed_data","N_BS_Con.rds")
+saveRDS(N_ECFU_Con_lm, file = N_ECFU_Con_location)
+
+N_ECFU_no3_location <- here::here("data","processed_data","N_BS_no3.rds")
+saveRDS(N_ECFU_no3_lm, file = N_ECFU_no3_location)
+
+N_ECFU_Turb_location <- here::here("data","processed_data","N_BS_Turb.rds")
+saveRDS(N_ECFU_Turb_lm, file = N_ECFU_Turb_location)
+```
+
+</br>
 ### In summary, exploration of correlation of stream health indicators showed that most interactions were non-significant.

### 2021-10-29 products/manuscript/Manuscript.Rmd
Project Part 3 - CC

@@
 ---
-
 # Summary/Abstract
 
+</br>
 # Introduction
-
 _When monitoring freshwater streams, stream health is typically measured by its physical characteristics (chemistry, surrounding land use area, channelization, etc.); Microbial (fecal indicator), and ecology (diversity of indicator macroinvertebrates).
@@
 Specifically, can you use one of the indicators mentioned (physical, microbial, and ecological) to describe both of the others to make a statement about overall stream health?_
-
+</br>
 _I would expect that because the all three are correlated with urbanization/anthropocentric impact, a negative impact in one indicator category will indicate a correlated negative impact in the other categories.
 The only caveat is that because level of noise in stream health data, significant correlations may be difficult to find._
-
+</br>
 # Methods and Results
-
 _UOWN has provided a great deal of data dating back to 2003.
@@
 Finally, I will test the interaction of all three via multivariate analysis._
-
+</br>
 ## Data aquisition
@@
 However, while cleaning I had to eliminate "po4.mgl" and "temperature" due to significant amounts of missing data._
-
+</br>
 #### Load import packages:
@@
 library(readxl)
+library(here)
+library(png)
+library(knitr)
 ```
-
+</br>
 #### Lodaing data and providing a summary for the raw data
@@
 data_location <- here::here("data/raw_data/UOWN_data_master_spring2021 (2).xlsx")
-
 rawdata <- readxl::read_excel(data_location)
-
 glimpse(rawdata)
 ```
+</br>
+## Data import and cleaning
+_Below is the cleaned data produced from our data processing. Data was processed using the processingscript.R file and includes the following steps:_
+_1. Load rawdata from excel file. Note the use of the here() package and not absolute paths_
+</br>
+_2. Designate NA's as actually missing values instead of factor data._
+</br>
+_3.
+Select variables of interest.
+These are variables I think are good indicators of stream health or are important identifying labels.
+Such variables are listed here: WSID, biological_score, conductivity.uscm, turbidity.ntu, po4.mgL, no3.mgL, pH, temperature.c,	e.coli.cfu, month, year,	and day.
+This was done using the Select() function._
+</br>
+_There is 3,075 rows and 12 columns...
+So, a lot of data.
+This is because data has been collected for ~20 years.
+Unfortunately, I am seeing a lot of NA's in the data.
+Because this is citizen science monitoring data, NA most likely means the data wasn't collected during that event and is a form of human error bias._
+</br>
+_4. Use NA assessment to see determine which variables to remove from assessment. This was done using the is_na function._
+_Several of the columns have more than half of the data missing.
+This is pretty common for citizen science data, seeing as collection is opportunistic.
+As such, I decided to remove P04 and temperature, because they are missing a lot of data.
+Finally, I decided to remove all missing values and see the condition of the data knowing that there are now no missing values.
+I removed NA data using the na.omit() function_
+</br>
+_5.
+Change all numeric columns to be read as numeric instead of character.
+This sometimes happens when R tries to read in data from an external source, but is an easy fix with the as.numeric function.
+This is important because my analysis uses linear regression functions and they are coded to only deal with numeric values._
+ </br>                          
+_6. Separate the data into three separate dataframes based on location ID: "MIDO", "NORO", and "BICO". This is to separate out any bias based on location of the streams within ACC._
+</br>
+_7.
+Because the number of rows between each site data frame and between years is not consistent, measuring between sites and through time may be biased.
+That being said, my original question refers to the correlation between different indicators of stream health, not how they change spatial or temporally.
+Therefore, these space and time inconsistencies won't affect the analyses._
+</br>
+_8. Saved cleand data files_
+```{r}
+#Set the location of the desired dataframes
+MIDO_location <- here::here("data","processed_data","MIDO.RDS")
+NORO_location <- here::here("data","processed_data","NORO.RDS")
+BICO_location <- here::here("data","processed_data","BICO.RDS")
 
-## Data import and cleaning
+#Load in the Dataframes
+MIDO <- readRDS(MIDO_location)
+NORO <- readRDS(NORO_location)
+BICO <- readRDS(BICO_location)
+```
+## Exploratory analysis
+_The following is an exploration of Stream Health Indicator correlation and comparison. The exploration will be structured as follows:_
+_Each dataframe (MIDO, NORO, and BICO) will be analyzed separately.
+Within each dataframe, there are data explaining biological indicator values, stream chemistry, and fecal indicators over time.
+The only group with multiple components is stream chemistry.
+I plan to do a series of correlative analysis showing if there is some sort of relationship between the three indicator types to help inform further analysis via principle component, multivariate analysis, etc.
+Therefore, the goal of this exploration is basically to summarize how these indicators correlate._
+### Below is the Linear Regression analysis from the Exploratory Data Analysis
+#### MIDO
+_Outcome: Biological Score_
+```{r}
+#Define data location
+MIDO_bs_lm_location <- here::here("results","MIDO_bs_lm.png")
+#Load in saved image from results folder
+include_graphics(MIDO_bs_lm_location)
+```
+_Outcome: E. coli (cfu)_
+```{r}
+#Define data location
+MIDO_ecfu_lm_location <- here::here("results","MIDO_ECFU_lm.png")
+#Load in the data
+include_graphics(MIDO_ecfu_lm_location)
+```
 
-## Exploratory analysis
+#### BICO
+_Outcome: Biological Score_
+```{r}
+#Define data location
+BICO_bs_lm_location <- here::here("results","BICO_bs_lm.png")
+#Load in saved image from results folder
+include_graphics(BICO_bs_lm_location)
+```
+_Outcome: E. coli (cfu)_
+```{r}
+#Define data location
+BICO_ecfu_lm_location <- here::here("results","BICO_ECFU_lm.png")
+#Load in the data
+include_graphics(BICO_ecfu_lm_location)
+```
 
+#### NORO
+_Outcome: Biological Score_
+```{r}
+#Define data location
+NORO_bs_lm_location <- here::here("results","NORO_bs_lm.png")
+#Load in saved image from results folder
+include_graphics(NORO_bs_lm_location)
+```
+_Outcome: E. coli (cfu)_
+```{r}
+#Define data location
+NORO_ecfu_lm_location <- here::here("results","NORO_ECFU_lm.png")
+#Load in the data
+include_graphics(NORO_ecfu_lm_location)
+```
 ## Full analysis
-
+_At this point,the main take-aways from data analysis thus far is that correlation of stream health indicators showed that most interactions were non-significant, with the exception of E.
+coli (cfu) and turbidity.
+Additionally, the data set is most likely not skewed by influential outliers._
+</br>
+_Moving forward, I want to focus in on just the MIDO dataset.
+As you can tell, working with many variables in three different data sets can be quite cumbersome to read through.
+For ease of interpretation, the MIDO data set will serve as the data frame where I test how robust the linear fit models actually are for the data.
+MIDO is the only data set large enough that when we split the data set into a "test" and "train" data set, I actually have confidence in the resulting model fit analysis.
+For comparison, MIDO has an n = 88; NORO has an n = 35; and BICO has an n = 23.
+The only thing we are missing by dropping the other sites is assessing site-by-site comparison, which I am choosing not to focus on anyway._
+</br>
+_The results indicated that only the relationship between E.
+Coli (cfu) and Conductivity were significant for both the train data (p = 1.374747e-05) and test data (p = 1.89214e-06).
+However, both models had high RMSE values, indicating that the models did not fit the data well and could not be used to predict E.
+coli (cfu) from conductivity, despite a significant relationship._
+</br>
 # Discussion

### 2021-11-03 products/manuscript/Manuscript.Rmd
Misc. edits - CC

@@
 ---
-# Summary/Abstract
 
-</br>
-# Introduction
-_When monitoring freshwater streams, stream health is typically measured by its physical characteristics (chemistry, surrounding land use area, channelization, etc.); Microbial (fecal indicator), and ecology (diversity of indicator macroinvertebrates).
-However, these categories are usually analyzed separately to assess stream health.
-As such, my question is how well are these parameters correlated with each other.
-Specifically, can you use one of the indicators mentioned (physical, microbial, and ecological) to describe both of the others to make a statement about overall stream health?_
-</br>
-_I would expect that because the all three are correlated with urbanization/anthropocentric impact, a negative impact in one indicator category will indicate a correlated negative impact in the other categories.
-The only caveat is that because level of noise in stream health data, significant correlations may be difficult to find._
-</br>
-# Methods and Results
-_UOWN has provided a great deal of data dating back to 2003.
-First,I will need to wrangle missing data values that are a result of inconsistent sampling efforts.
-This will make the data easier to work with and narrow down a time frame for the data that is consistent and free of sampling bias.
-Next, I will analyze and compare trends in the change of chemical (conductivity, turbidity,	no3.mgL); ecological(Biological Score); and microbial indicator data (e.coli CFU).
-Generally speaking, increases in chemical properties like conductivity, turbidity,	no3.mgL and microbial indicator abundance (i.e.
-ecoli CFU) indicate decreased stream health.
-While a decrease in biological score indicates decrease in stream health.
-Next, since biological scores are used as a measure of broader ecological health, I will run t-tests on if the other factors significantly effect biological score.
-Finally, I will test the interaction of all three via multivariate analysis._
-</br>
-## Data aquisition
-_The data I am using is the publically avaialable data provided by the Upper Oconee Watershed Network (UOWN).
-This data is part of a citizen science training effort for routine watershed-scale monitoring of the upper oconee river watershed.
-Trained freshwater scientists assist in training members of the community to collect both chemical and biological stream characteristics that are useful in describing stream health.
-All data is compliled and posted in raw form on the UOWN website.
-Of interest to me are the following variables: WSID (ID number),	biological_score,	conductivity, turbidity,	no3.mgL,	e.coli.cfu,	month,	year,	day.
-The raw data shows 3,075 observations before cleaning.
-However, while cleaning I had to eliminate "po4.mgl" and "temperature" due to significant amounts of missing data._
-</br>
-#### Load import packages:
-```{r}
+```{r, echo=FALSE, results='hide', message=FALSE}
+#Load important packages:
 library(tidyverse)
@@
 ```
+
 </br>
-#### Lodaing data and providing a summary for the raw data
-```{r, warning=FALSE}
+
+```{r, warning=FALSE, echo=FALSE, results='hide'}
+#Lodaing data and providing a summary for the raw data
 data_location <- here::here("data/raw_data/UOWN_data_master_spring2021 (2).xlsx")
@@
 ```
-</br>
-## Data import and cleaning
-_Below is the cleaned data produced from our data processing. Data was processed using the processingscript.R file and includes the following steps:_
-_1. Load rawdata from excel file. Note the use of the here() package and not absolute paths_
-</br>
-_2. Designate NA's as actually missing values instead of factor data._
-</br>
-_3.
-Select variables of interest.
-These are variables I think are good indicators of stream health or are important identifying labels.
-Such variables are listed here: WSID, biological_score, conductivity.uscm, turbidity.ntu, po4.mgL, no3.mgL, pH, temperature.c,	e.coli.cfu, month, year,	and day.
-This was done using the Select() function._
-</br>
-_There is 3,075 rows and 12 columns...
-So, a lot of data.
-This is because data has been collected for ~20 years.
-Unfortunately, I am seeing a lot of NA's in the data.
-Because this is citizen science monitoring data, NA most likely means the data wasn't collected during that event and is a form of human error bias._
-</br>
-_4. Use NA assessment to see determine which variables to remove from assessment. This was done using the is_na function._
-_Several of the columns have more than half of the data missing.
-This is pretty common for citizen science data, seeing as collection is opportunistic.
-As such, I decided to remove P04 and temperature, because they are missing a lot of data.
-Finally, I decided to remove all missing values and see the condition of the data knowing that there are now no missing values.
-I removed NA data using the na.omit() function_
-</br>
-_5.
-Change all numeric columns to be read as numeric instead of character.
-This sometimes happens when R tries to read in data from an external source, but is an easy fix with the as.numeric function.
-This is important because my analysis uses linear regression functions and they are coded to only deal with numeric values._
- </br>                          
-_6. Separate the data into three separate dataframes based on location ID: "MIDO", "NORO", and "BICO". This is to separate out any bias based on location of the streams within ACC._
-</br>
-_7.
-Because the number of rows between each site data frame and between years is not consistent, measuring between sites and through time may be biased.
-That being said, my original question refers to the correlation between different indicators of stream health, not how they change spatial or temporally.
-Therefore, these space and time inconsistencies won't affect the analyses._
-</br>
-_8. Saved cleand data files_
-```{r}
+
+```{r, echo=FALSE, results='hide'}
 #Set the location of the desired dataframes
@@
 ```
+
+# Summary/Abstract
+
+</br>
+
+# Introduction
+
+_When monitoring freshwater streams, stream health is typically measured by its physical characteristics (chemistry, surrounding land use area, channelization, etc.); Microbial (fecal indicator), and ecology (diversity of indicator macroinvertebrates).
+However, these categories are usually analyzed separately to assess stream health.
+As such, my question is how well are these parameters correlated with each other.
+Specifically, can you use one of the indicators mentioned (physical, microbial, and ecological) to describe both of the others to make a statement about overall stream health?_
+
+</br>
+
+_I would expect that because the all three are correlated with urbanization/anthropocentric impact, a negative impact in one indicator category will indicate a correlated negative impact in the other categories.
+The only caveat is that because level of noise in stream health data, significant correlations may be difficult to find._
+
+</br>
+
+# Methods and Results
+
+_UOWN has provided a great deal of data dating back to 2003.
+First,I will need to wrangle missing data values that are a result of inconsistent sampling efforts.
+This will make the data easier to work with and narrow down a time frame for the data that is consistent and free of sampling bias.
+Next, I will analyze and compare trends in the change of chemical (conductivity, turbidity,	no3.mgL); ecological(Biological Score); and microbial indicator data (e.coli CFU).
+Generally speaking, increases in chemical properties like conductivity, turbidity,	no3.mgL and microbial indicator abundance (i.e.
+ecoli CFU) indicate decreased stream health.
+While a decrease in biological score indicates decrease in stream health.
+Next, since biological scores are used as a measure of broader ecological health, I will run t-tests on if the other factors significantly effect biological score.
+Finally, I will test the interaction of all three via multivariate analysis._
+
+</br>
+
+## Data aquisition
+
+_The data I am using is the publically avaialable data provided by the Upper Oconee Watershed Network (UOWN).
+This data is part of a citizen science training effort for routine watershed-scale monitoring of the upper oconee river watershed.
+Trained freshwater scientists assist in training members of the community to collect both chemical and biological stream characteristics that are useful in describing stream health.
+All data is compliled and posted in raw form on the UOWN website.
+Of interest to me are the following variables: WSID (ID number),	biological_score,	conductivity, turbidity,	no3.mgL,	e.coli.cfu,	month,	year,	day.
+The raw data shows 3,075 observations before cleaning.
+However, while cleaning I had to eliminate "po4.mgl" and "temperature" due to significant amounts of missing data._
+
+</br>
+
+## Data import and cleaning
+
+_Below is the cleaned data produced from our data processing. Data was processed using the processingscript.R file and includes the following steps:_
+_1. Load rawdata from excel file. Note the use of the here() package and not absolute paths_
+
+</br>
+
+_2. Designate NA's as actually missing values instead of factor data._
+
+</br>
+
+_3.
+Select variables of interest.
+These are variables I think are good indicators of stream health or are important identifying labels.
+Such variables are listed here: WSID, biological_score, conductivity.uscm, turbidity.ntu, po4.mgL, no3.mgL, pH, temperature.c,	e.coli.cfu, month, year,	and day.
+This was done using the Select() function._
+
+</br>
+
+_There is 3,075 rows and 12 columns...
+So, a lot of data.
+This is because data has been collected for ~20 years.
+Unfortunately, I am seeing a lot of NA's in the data.
+Because this is citizen science monitoring data, NA most likely means the data wasn't collected during that event and is a form of human error bias._
+
+</br>
+
+_4. Use NA assessment to see determine which variables to remove from assessment. This was done using the is_na function._
+_Several of the columns have more than half of the data missing.
+This is pretty common for citizen science data, seeing as collection is opportunistic.
+As such, I decided to remove P04 and temperature, because they are missing a lot of data.
+Finally, I decided to remove all missing values and see the condition of the data knowing that there are now no missing values.
+I removed NA data using the na.omit() function_
+
+</br>
+
+_5.
+Change all numeric columns to be read as numeric instead of character.
+This sometimes happens when R tries to read in data from an external source, but is an easy fix with the as.numeric function.
+This is important because my analysis uses linear regression functions and they are coded to only deal with numeric values._
+
+ </br>
+
+_6. Separate the data into three separate dataframes based on location ID: "MIDO", "NORO", and "BICO". This is to separate out any bias based on location of the streams within ACC._
+
+</br>
+
+_7.
+Because the number of rows between each site data frame and between years is not consistent, measuring between sites and through time may be biased.
+That being said, my original question refers to the correlation between different indicators of stream health, not how they change spatial or temporally.
+Therefore, these space and time inconsistencies won't affect the analyses._
+
+</br>
+
 ## Exploratory analysis
+
 _The following is an exploration of Stream Health Indicator correlation and comparison. The exploration will be structured as follows:_
@@
 Therefore, the goal of this exploration is basically to summarize how these indicators correlate._
+
 ### Below is the Linear Regression analysis from the Exploratory Data Analysis
+
 #### MIDO
+
 _Outcome: Biological Score_
-```{r}
+```{r, echo=FALSE}
 #Define data location
@@
 _Outcome: E. coli (cfu)_
-```{r}
+```{r, echo=FALSE}
 #Define data location
@@
 #### BICO
+
 _Outcome: Biological Score_
-```{r}
+```{r, echo=FALSE}
 #Define data location
@@
 #### NORO
+
 _Outcome: Biological Score_
-```{r}
+```{r, echo=FALSE}
 #Define data location
@@
 ```
+</br>
+
 ## Full analysis
+
 _At this point,the main take-aways from data analysis thus far is that correlation of stream health indicators showed that most interactions were non-significant, with the exception of E.
@@
 Additionally, the data set is most likely not skewed by influential outliers._
+
 </br>
+
 _Moving forward, I want to focus in on just the MIDO dataset.
@@
 The only thing we are missing by dropping the other sites is assessing site-by-site comparison, which I am choosing not to focus on anyway._
+
 </br>
+
 _The results indicated that only the relationship between E.
@@
 coli (cfu) from conductivity, despite a significant relationship._
+
 </br>
+
 # Discussion

### 2021-11-12 code/analysis_code/analysis_uown.Rmd
LASSO_Model - CC

@@
 ```{r}
-library(tidyverse)
-library(tidymodels)
-library(here)
+library(tidyverse) #Working with multiple Tidy packages
+library(tidymodels) #Building models
+library(dplyr) #data manipulation
+library(here) #setting pathways for saving files
+library(rpart) #Model fitting
+library(ranger) #Model fitting
+library(glmnet) #Model fitting
 ```
@@
 ```
+
+# Part 3: LASSO Modeling for Ideal Combination of Stream Health Indicators to Predict Biological SCore and E. Coli Colony Forming Units in MIDO Data.
+
+# Biological Score
+
+## Data Set Up:
+
+_Take ID and time variables out of MIDO Data frame. This will help reduce errors when we try to run the Model._
+```{r}
+MIDO_LASSO <- MIDO %>%
+  select(-"WSID", -"stream_ID", -"month", -"year", -"day", -"pH")
+```
+
+</br>
+
+_log transform E. coli data to fix bias in E. coli data_
+```{r}
+MIDO_LASSO <- MIDO_LASSO %>%
+  mutate(e.coli.cfu = log10(MIDO_LASSO$e.coli.cfu))
+```
+
+</br>
+
+_Set seed: This sets a random number generator with initial (pseudo)random values set as "123". We will need a series of random numbers created for our machine learning analysis._
+```{r}
+#Set random number generator.
+set.seed(123)
+```
+
+ _Test/Train data split: Split data into a training data set and a test data set.
+The training set will contain most of the original data (70%), while the test data set will have a smaller portion (30%).
+The training data is used to fit a model and the test data is to assess how good of a fit the data is._
+
+</br>
+
+## Null Model For MIDO Data
+
+_5-fold cross validation, 5 times repeated for MIDO data: Here we are setting a cross-validation of the machine learning models.
+Cross-validation is used to measure how the results of our machine learning models will generalize to an independent data set.
+As such, the folds created will be be 5 random sub-samples of the train data set to test the validity of our models within the train data set.
+The 5x5 structure is arbitrary._
+```{r}
+#Create folds for LASSO Cross Validation.
+fold_BS <- vfold_cv(MIDO_LASSO, v = 5, repeats = 5, strata = biological_score)
+```
+
+_Creating the recipe for Biological Score vs all predictors_
+
+```{r}
+BS.recipe <- 
+  recipe(biological_score ~ ., data = MIDO_LASSO) %>%
+  step_zv(all_predictors()) %>%
+  step_normalize(all_predictors())
+
+BS.recipe
+```
+_Setting linear regression model to assess relationship between Biological Score (outcome) and all other predictor variables._
+```{r}
+lm_mod <- linear_reg() %>% set_engine("lm") %>% set_mode("regression")
+```
+
+_However, first we need to create our null model to test against._
+
+</br>
+
+### Null Model:
+
+ </br>
+
+_Creates null model recipe. When we call this term, it will indicate in our workflow that Biological Score will be predicted by a value of 1 (NULL)._ 
+```{r}
+Null_recipe_lm <- recipe(biological_score ~ 1, data = MIDO_LASSO)
+```
+
+_Creating the Workflow: this creates a set workflow for running a null linear regression model with body temperature as the outcome._
+```{r}
+null_wf <- workflow() %>% add_model(lm_mod) %>% add_recipe(Null_recipe_lm)
+```
+
+_Here, I am going to fit the null model created in the above workflow to the folds made from the train data set._
+```{r}
+null_lm <- fit_resamples(null_wf, resamples = fold_BS)
+```
+_Calculate RMSE for the train data linear model._
+```{r}
+Null_Met <- collect_metrics(null_lm)
+
+Null_Met
+```
+_RMSE = 6.63, with a standard deviation of 0.13. This will serve as our check to test our models against latter on.
+
+</br>
+
+## LASSO Model
+
+</br>
+
+_Specifying The Model: LASSO_
+```{r}
+lasso_mod <- 
+  linear_reg(penalty = tune(), mixture = 1) %>% 
+  set_engine("glmnet")
+```
+
+</br>
+
+_Creating a Workflow: LASSO_
+```{r}
+lasso_wf <- workflow() %>%
+  add_model(lasso_mod) %>%
+  add_recipe(BS.recipe)
+```
+
+</br>
+
+_Create Tuning Grid: LASSO_
+```{r}
+LASSO_grid <- grid_regular(penalty(), levels = 50)
+```
+
+</br>
+
+_Cross Validation with tune_grid(): LASSO_
+```{r}
+lasso_resample <- 
+  lasso_wf %>%
+  tune_grid(resamples = fold_BS,
+            grid = LASSO_grid,
+            control = control_grid(verbose = FALSE, save_pred = TRUE),
+            metrics = metric_set(rmse))
+
+lasso_resample %>%
+  collect_metrics()
+```
+
+</br>
+
+_Plot model performance using autoplot()_
+```{r}
+#Plot of MIDO data LASSO model performance
+lasso_resample %>%
+  autoplot()
+```
+
+_Showing and selecting best performing Models_
+```{r}
+#Showing best performing LASSO models
+lasso_resample %>%
+  show_best()
+
+#Selects best performing model
+best_lasso <- lasso_resample %>%
+  select_best()
+```
+
+_This shows that model 47 is the best performing models (RMSE = 6.50; STE = 0.16). However, it doesn't really perform any better than the null model, making it a bad fit to the data._
+
+</br>
+
+_Creating final fit based on best model permutation and plotting predicted values from that final fit model_
+```{r}
+lasso_final_wf <- 
+  lasso_wf %>% 
+  finalize_workflow(best_lasso)
+
+lasso_final_wf
+
+#Create workflow for fitting model to train_data2 predictions
+lasso_final_fit <- 
+  lasso_final_wf %>%
+  fit(MIDO_LASSO) 
+```
+
+_Calculating residuals:_
+```{r}
+#Manually calculate residuals for MIDO data LASSO models between real and predicted values.
+lasso_residuals <- lasso_final_fit %>%
+  augment(MIDO_LASSO) %>% #use augment() to make predictions from train data
+  select(c(.pred, biological_score)) %>%
+  mutate(.resid = biological_score - .pred) #calculate residuals and make new row.
+
+lasso_residuals
+```
+_model predictions from tuned model vs actual outcomes_
+```{r}
+lasso_pred_plot <- ggplot(lasso_residuals, 
+                          aes(x = biological_score, 
+                              y = .pred)) + 
+  geom_point() + 
+  labs(title = "Predictions vs Actual: LASSO", 
+       x = "Biological Score Outcome", 
+       y = "Biological Score Prediction")
+lasso_pred_plot
+```
+_plot residuals vs predictions_
+```{r}
+lasso_residual_plot <- ggplot(lasso_residuals, 
+                              aes(y = .resid, 
+                                  x = .pred)) + 
+  geom_point() + 
+  labs(title = "Predictions vs Residuals: LASSO", 
+       x = "Biological Score Prediction", 
+       y = "Residuals")
+plot(lasso_residual_plot) #view plot
+```
+
+# LASSO: E. coli cfu
+
+## Data Set Up:
+
+_Set seed: This sets a random number generator with initial (pseudo)random values set as "123". We will need a series of random numbers created for our machine learning analysis._
+```{r}
+#Set random number generator.
+set.seed(123)
+```
+
+ _Test/Train data split: Split data into a training data set and a test data set.
+The training set will contain most of the original data (70%), while the test data set will have a smaller portion (30%).
+The training data is used to fit a model and the test data is to assess how good of a fit the data is._
+
+</br>
+
+## Null Model For MIDO Data
+
+_5-fold cross validation, 5 times repeated for MIDO data: Here we are setting a cross-validation of the machine learning models.
+Cross-validation is used to measure how the results of our machine learning models will generalize to an independent data set.
+As such, the folds created will be be 5 random sub-samples of the train data set to test the validity of our models within the train data set.
+The 5x5 structure is arbitrary._
+```{r}
+#Create folds for LASSO Cross Validation.
+fold_EC <- vfold_cv(MIDO_LASSO, v = 5, repeats = 5, strata = e.coli.cfu)
+```
+
+_Creating the recipe for Biological Score vs all predictors_
+
+```{r}
+EC.recipe <- 
+  recipe(e.coli.cfu ~ ., data = MIDO_LASSO) %>%
+  step_zv(all_predictors()) %>%
+  step_normalize(all_predictors())
+
+EC.recipe
+```
+_Setting linear regression model to assess relationship between Biological Score (outcome) and all other predictor variables._
+```{r}
+lm_mod <- linear_reg() %>% set_engine("lm") %>% set_mode("regression")
+```
+
+_However, first we need to create our null model to test against._
+
+</br>
+
+### Null Model:
+
+ </br>
+
+_Creates null model recipe. When we call this term, it will indicate in our workflow that E. Coli cfu will be predicted by a value of 1 (NULL)._ 
+```{r}
+Null_recipe_lm_EC <- recipe(e.coli.cfu ~ 1, data = MIDO_LASSO)
+```
+
+_Creating the Workflow: this creates a set workflow for running a null linear regression model with E. coli cfu as the outcome._
+```{r}
+null_wf_EC <- workflow() %>% add_model(lm_mod) %>% add_recipe(Null_recipe_lm_EC)
+```
+
+_Here, I am going to fit the null model created in the above workflow to the folds made from the train data set._
+```{r}
+null_lm_EC <- fit_resamples(null_wf_EC, resamples = fold_EC)
+```
+
+</br>
+
+_Calculate RMSE for the train data linear model._
+```{r}
+Null_Met_EC <- collect_metrics(null_lm_EC)
+
+Null_Met_EC
+```
+_RMSE = 0.52, with a standard deviation of 0.014. This will serve as our check to test our models against latter on._
+
+</br>
+
+## LASSO Model
+
+</br>
+
+_Specifying The Model: LASSO_
+```{r}
+lasso_mod <- 
+  linear_reg(penalty = tune(), mixture = 1) %>% 
+  set_engine("glmnet")
+```
+
+</br>
+
+_Creating a Workflow: LASSO_
+```{r}
+lasso_wf_EC <- workflow() %>%
+  add_model(lasso_mod) %>%
+  add_recipe(EC.recipe)
+```
+
+</br>
+
+_Create Tuning Grid: LASSO_
+```{r}
+LASSO_grid <- grid_regular(penalty(), levels = 50)
+```
+
+</br>
+
+_Cross Validation with tune_grid(): LASSO_
+```{r}
+lasso_resample_EC <- 
+  lasso_wf_EC %>%
+  tune_grid(resamples = fold_EC,
+            grid = LASSO_grid,
+            control = control_grid(verbose = FALSE, save_pred = TRUE),
+            metrics = metric_set(rmse))
+
+lasso_resample_EC %>%
+  collect_metrics()
+```
+
+</br>
+
+_Plot model performance using autoplot()_
+```{r}
+#Plot of MIDO data LASSO model performance
+lasso_resample_EC %>%
+  autoplot()
+```
+
+_Showing and selecting best performing Models_
+```{r}
+#Showing best performing LASSO models
+lasso_resample_EC %>%
+  show_best()
+
+#Selects best performing model
+best_lasso_EC <- lasso_resample_EC %>%
+  select_best()
+```
+
+_This shows that model 41 is the best performing models (RMSE = 0.50; STE = 0.012). However, it doesn't really perform any better than the null model, making it a bad fit to the data._
+
+</br>
+
+_Creating final fit based on best model permutation and plotting predicted values from that final fit model_
+```{r}
+lasso_final_wf_EC <- 
+  lasso_wf_EC %>% 
+  finalize_workflow(best_lasso_EC)
+
+lasso_final_wf_EC
+
+#Create workflow for fitting model to train_data2 predictions
+lasso_final_fit_EC <- 
+  lasso_final_wf_EC %>%
+  fit(MIDO_LASSO) 
+```
+
+_Calculating residuals:_
+```{r}
+#Manually calculate residuals for MIDO data LASSO models between real and predicted values.
+lasso_residuals_EC <- lasso_final_fit_EC %>%
+  augment(MIDO_LASSO) %>% #use augment() to make predictions from train data
+  select(c(.pred, e.coli.cfu)) %>%
+  mutate(.resid = e.coli.cfu - .pred) #calculate residuals and make new row.
+
+lasso_residuals_EC
+```
+_model predictions from tuned model vs actual outcomes_
+```{r}
+lasso_pred_plot_EC <- ggplot(lasso_residuals_EC, 
+                          aes(x = e.coli.cfu, 
+                              y = .pred)) + 
+  geom_point() + 
+  labs(title = "Predictions vs Actual: LASSO", 
+       x = "E. coli cfu Outcome", 
+       y = "E. coli cfu Prediction")
+lasso_pred_plot_EC
+```
+_plot residuals vs predictions_
+```{r}
+lasso_residual_plot_EC <- ggplot(lasso_residuals_EC, 
+                              aes(y = .resid, 
+                                  x = .pred)) + 
+  geom_point() + 
+  labs(title = "Predictions vs Residuals: LASSO", 
+       x = "E. coli cfu Prediction", 
+       y = "Residuals")
+plot(lasso_residual_plot_EC) #view plot
+```

### 2021-11-12 code/analysis_code/analysis_uown.Rmd
Project_p4 - CC

@@
 </br>
+
 ### Load all the created linear models from the processing data markdown.
@@
 This is because the cooksdistance() function calls for a linear model input.
+
+</br>
+
 #### MIDO
+
+</br>
+
 _Biological Score_
@@
 ```
+
+</br>
+
 _ECFU_
@@
 
+</br>
+
 #### BICO
+
+</br>
+
 _Biological Score_
@@
 
+</br>
+
 #### NORO
+
+</br>
+
 _Biological Score_
@@
 
+</br>
+
 ### Outlier identification using Cook's Distance.
+
+</br>
+
 _Outliers in these data could be potentially ttreated as errors in reporting.
@@
 Therefore, an argument can be made that assessing for anomalies in the data could help tease apart the true relationships that exist between different stream health indicators._
-</br>
+
+</br>
+
 _As such, we will start by analyzing the data for outliers by evaluating relationships with cook's distance._
-</br>
+
+</br>
+
 #### Set sample size for sample size parameter (4/n) for each data set (MIDO, NORO, BICO)
@@
 ```
-</br>
+
+</br>
+
 #### MIDO
-</br>
+
+</br>
+
 _Outcome: Biological Score_
@@
 </br>
+
 #### NORO
+
 </br>
@@
 </br>
+
 _Outcome: E.coli CFU_
@@
 </br>
+
 #### BICO
-</br>
+
+</br>
+
 _Outcome: Biological Score_
@@
 </br>
+
 _The results of the Cook's Distance show that while there are outliers, very few of them actually exceed a Cook's Distance value (CDV) of 1.
@@
 As such, it is my educated opinion to leave all data points in, as I do not believe they are skewing liner modle results in a significant enough way to warrant removal._
+
 </br>
@@
 # Part 2: Linear Model Fit Assessment for MIDO Data
-</br>
+
+</br>
+
 _At this point,the main take-aways from data analysis thus far is that correlation of stream health indicators showed that most interactions were non-significant, with the exception of E.
@@
 Additionally, the data set is most likely not skewed by influential outliers._
-</br>
+
+</br>
+
 _Moving forward, I want to focus in on just the MIDO dataset.
@@
 The only thing we are missing by dropping the other sites is assessing site-by-site comparison, which I am choosing not to focus on anyway._
-</br>
+
+</br>
+
 #### Splitting MIDO Data
@@
 ```
-</br>
+
+</br>
+
 _Create a recipe for each linear model relationship in the MIDO dataframe._
@@
 ```
+
+</br>
+
 _With our seven linear model recipes created, we will define of linear regression model pipe._
@@
 ```
+
+</br>
+
 _Next, create a workflow that adds our seven recipes and model together._
@@
 
+</br>
+
 ## Modeling Train Data Predictions Using Workflows
 _Using the workflow above, lets now fit each pf the seven stream health indicator linear models to the train data set._
-</br>
+
+</br>
+
 _Defining a command that runs model fitting to each of the seven models_
@@
 ```
-</br>
+
+</br>
+
 _Pull linear regression fit models using parsnip()._
@@
 </br>
+
 #### Using Workflows to Make Predictions in the train data set.
@@
 ```
-</br>
+
+</br>
+
 #### Using Root Mean Squared Error (RMSE) to Assess Model Fit.
+
+</br>
 
@@
 </br>
+
 #####Analysis here#####
-</br>
+
+</br>
+
 ## Modeling Test Data Predictions Using Workflows
 _Using the workflow above, lets now fit each pf the seven stream health indicator linear models to the test data set._
-</br>
+
+</br>
+
 _Defining a command that runs model fitting to each of the seven models_
@@
 #####Analysis here#####
-</br>
+
+</br>
+
 #### Using Workflows to Make Predictions in the test data set.
@@
 ```
-</br>
+
+</br>
+
 #### Using Root Mean Squared Error (RMSE) to Assess Model Fit.
@@
 ```
+</br>
+
 #### Save significant Test and Train linear models to the results folder.
@@
 ```
-</br>
+
+</br>
+
 #### Save significant Test and Train RMSE to the results folder.
@@
 ```
-
- _Test/Train data split: Split data into a training data set and a test data set.
-The training set will contain most of the original data (70%), while the test data set will have a smaller portion (30%).
-The training data is used to fit a model and the test data is to assess how good of a fit the data is._
 
@@
 
-_Creating the Workflow: this creates a set workflow for running a null linear regression model with body temperature as the outcome._
+_Creating the Workflow: this creates a set workflow for running a null linear regression model with biological score as the outcome._
 ```{r}
@@
 ```
-_Calculate RMSE for the train data linear model._
+_Calculate RMSE for the Null linear model._
 ```{r}
@@
 ```
-_RMSE = 6.63, with a standard deviation of 0.13. This will serve as our check to test our models against latter on.
+_RMSE = 6.63, with a standard deviation of 0.13. This will serve as our check to test our models against latter on._
 
@@
 lasso_resample %>%
-  autoplot()
+  autoplot(main="MIDO Data LASSO Performance")
+
 ```
@@
 
- _Test/Train data split: Split data into a training data set and a test data set.
-The training set will contain most of the original data (70%), while the test data set will have a smaller portion (30%).
-The training data is used to fit a model and the test data is to assess how good of a fit the data is._
-
 </br>
@@
 
-_Creating the recipe for Biological Score vs all predictors_
+_Creating the recipe for E. coli cfu vs all predictors_
 
@@
 ```
-_Setting linear regression model to assess relationship between Biological Score (outcome) and all other predictor variables._
+_Setting linear regression model to assess relationship between E. coli cfu (outcome) and all other predictor variables._
 ```{r}

### 2021-11-12 code/processing_code/Data_Exploration.Rmd
Project_p4 - CC

@@
 ```{r}
+#Creates locations from file pathways that the required data frame .RDS files are stored.
 MIDO_location <- here::here("data","processed_data","MIDO.RDS")
@@
 
+#Reads files called from defined locations
 MIDO <- readRDS(MIDO_location)
@@
 ```{r}
+#Plot biological score vs. conductivity
 BS_Con <- MIDO %>% 
@@
 ```{r}
+#Linear regression analysis for biological score versus conductivity.
 BS_Con_lm <- lm(biological_score ~ conductivity.uscm, data=MIDO)
@@
 ```{r}
+#Plot biological score vs. turbidity
 BS_Turb <- MIDO %>%
@@
 ### *E. Coli*
+
 </br>
+
 #### E. Coli Vs Conductivity

### 2021-11-12 products/manuscript/Manuscript.Rmd
Project_p4 - CC

@@
 ---
-
-```{r, echo=FALSE, results='hide', message=FALSE}
-#Load important packages:
-library(tidyverse)
-library(readxl)
-library(here)
-library(png)
-library(knitr)
-```
-
-</br>
-
-```{r, warning=FALSE, echo=FALSE, results='hide'}
-#Lodaing data and providing a summary for the raw data
-data_location <- here::here("data/raw_data/UOWN_data_master_spring2021 (2).xlsx")
-rawdata <- readxl::read_excel(data_location)
-glimpse(rawdata)
-```
-
-```{r, echo=FALSE, results='hide'}
-#Set the location of the desired dataframes
-MIDO_location <- here::here("data","processed_data","MIDO.RDS")
-NORO_location <- here::here("data","processed_data","NORO.RDS")
-BICO_location <- here::here("data","processed_data","BICO.RDS")
-
-#Load in the Dataframes
-MIDO <- readRDS(MIDO_location)
-NORO <- readRDS(NORO_location)
-BICO <- readRDS(BICO_location)
-```
 
@@
 
-### Below is the Linear Regression analysis from the Exploratory Data Analysis
-
-#### MIDO
-
-_Outcome: Biological Score_
-```{r, echo=FALSE}
-#Define data location
-MIDO_bs_lm_location <- here::here("results","MIDO_bs_lm.png")
-#Load in saved image from results folder
-include_graphics(MIDO_bs_lm_location)
-```
-_Outcome: E. coli (cfu)_
-```{r, echo=FALSE}
-#Define data location
-MIDO_ecfu_lm_location <- here::here("results","MIDO_ECFU_lm.png")
-#Load in the data
-include_graphics(MIDO_ecfu_lm_location)
-```
-
-#### BICO
-
-_Outcome: Biological Score_
-```{r, echo=FALSE}
-#Define data location
-BICO_bs_lm_location <- here::here("results","BICO_bs_lm.png")
-#Load in saved image from results folder
-include_graphics(BICO_bs_lm_location)
-```
-_Outcome: E. coli (cfu)_
-```{r, echo=FALSE}
-#Define data location
-BICO_ecfu_lm_location <- here::here("results","BICO_ECFU_lm.png")
-#Load in the data
-include_graphics(BICO_ecfu_lm_location)
-```
-
-#### NORO
-
-_Outcome: Biological Score_
-```{r, echo=FALSE}
-#Define data location
-NORO_bs_lm_location <- here::here("results","NORO_bs_lm.png")
-#Load in saved image from results folder
-include_graphics(NORO_bs_lm_location)
-```
-_Outcome: E. coli (cfu)_
-```{r, echo=FALSE}
-#Define data location
-NORO_ecfu_lm_location <- here::here("results","NORO_ECFU_lm.png")
-#Load in the data
-include_graphics(NORO_ecfu_lm_location)
-```
-</br>
-
 ## Full analysis
@@
 
-_Moving forward, I want to focus in on just the MIDO dataset.
+_Following correlation analysis during exploration, focus was shifted to just the MIDO dataset.
 As you can tell, working with many variables in three different data sets can be quite cumbersome to read through.
@@
 
+_Finally, LASSO analyses were performed with either biological score or E.
+coli cfu as the outcome of choice.
+The LASSO model analyzed which combination of stream health descriptive variables are the best for predicting either biological score or E.
+coli cfu values for the MIDO site through time._
+
+</br>
+
 _The results indicated that only the relationship between E.
@@
 However, both models had high RMSE values, indicating that the models did not fit the data well and could not be used to predict E.
-coli (cfu) from conductivity, despite a significant relationship._
+coli (cfu) from conductivity, despite a significant relationship.
+To test how well the LASSO models performed and how well they fit our data, Root Mean Squared Error (RMSE) was calculated for each model produced and for predicted values of MIDO stream health variables, using the LASSO models.
+Residuals between actual and predicted values were calculated to show how well the models could predict values.
+Finally, all RMSE values (actual and predicted data models) were compared to a null model for both biological score and E.
+coli cfu to evaluate model performance._
+
+</br>
+
+_The null linear models for biological score and E.
+coli cfu produced a RMSE = 6.63, with a standard deviation of 0.13 and RMSE = 0.52, with a standard deviation of 0.014, respectively.
+Results from the biological score LASSO machine learning model indicated that model 47 is the best performing models (RMSE = 6.50; STE = 0.16).
+However, based on comparative RMSE scores, model 47 does not really perform any better than the null model.
+This indicates that it is a bad fit to the MIDO data.
+Results from the E.
+coli cfu LASSO machine learning model indicated that model 41 is the best performing models (RMSE = 0.50; STE = 0.012).
+However, based on comparative RMSE scores, model 41 does not really perform any better than the null model.
+This indicates that it is a bad fit to the MIDO data._
 
@@
 
-## Summary and Interpretation
+</br>
 
-## Strengths and Limitations
+_Based on the linear modeling exploratory analysis and LASSO modeling for best combination of predictors, the results indicate that combinations of the variables biological score, E.
+coli cfu, nitrate concentration, conductivity, and turbidity are not able to predict values for biological score or E.
+coli cfu.
+This indicates that one cannot predict one measure of stream health from another.
+This is seen by a lack of consistent significant correlations between each stream health predictor and either biological score or E.
+coli cfu as well as LASSO modeling to find the ideal combination of stream health variables to predict biological score or E.
+coli cfu.
+Given these results, it is likely that the data was biased due to the  inconsistent and opportunistic sampling efforts done by citizen science [find citation for lack of reliability for citizen science data].
+When examining the data set, many values were observed to be missing, with a disproportionate amount of extent observations in what I suspect to be the more easily collected stream health variables.
+For example, significantly more NA observations were included in the biological score column relative to variables like turbidty and conductivity.
+This is because both turbidity and conductivity can be measured in situ with a meter in a matter of minutes, while biological score is calculated following collection and identification of macroinvertebrate families from samples [find reference for biological score assessment methodology].
+Therefore, biological score is much harder to measure.
+Additionally, E.
+coli cfu is processed by culturing E.
+coli from samples and generally requires a trained scientist to do [find reference for E.
+coli cfu qunatification methods].
+While confidence in values may be greater in this case, there is know a shortage of labor due to need for a specialized skill set to process and quantify E.
+coli cfu values._
+
+</br>
 
@@
 
+</br>
+
+_Based the results of this study, I would recommend that existing methodology for UOWN stream health characterization and monitoring be re-evaluated.
+The issue being that there seems to be a disparity in sampling effort, both in quality of samples and time spent sampling, for each different stream health metric.
+The result being that the term "healthy stream" does not have a unified definition in terms of the metrics used to quantify stream health.
+I would suggest future scientific studies be conducted to assess to relationship between stream health metrics in order to be able to predict metrics like biological score and E.
+coli cfu so citizen based monitoring does not have to collect these data while still having easier to measure variables that can be used to describe stream health in terms of biological score and E.
+coli cfu._
+
 # References

### 2021-11-12 results/Supplementary_Data.Rmd
Part4_edits - CC

---
title: "Supplementary Data"
author: "C. Coleman"
date: "11/12/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

### 2021-11-12 results/Supplementary_Data.Rmd
Supplementary_Figures - CC

@@
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+_The following markdown serves as the supplementary materials for the UOWN Stream Health Indicators Project.
+Included materials are all figures relevant to my manuscript and were derived from the analysis code._
+
+</br>
+
+_Load needed libraries_
+```{r}
+library(knitr)    # For knitting document and include_graphics function
+library(ggplot2)  # For plotting
+library(png)      # For grabbing the dimensions of png files
 ```
 
-## R Markdown
+</br>
 
-This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.
+_Figure 1: Plots of biological score versus conductivity, E. coli cfu, nitrate concentration, and turbidity for the MIDO site._
+```{r}
+bs_MIDO_location <- here::here("results", "MIDO_bs_lm.png")
 
-When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.
-You can embed an R code chunk like this:
-
-```{r cars}
-summary(cars)
+include_graphics(bs_MIDO_location)
 ```
 
-## Including Plots
+</br>
 
-You can also embed plots, for example:
+_Figure 2: Plots of E. coli cfu versus conductivity, nitrate concentration, and turbidity for the MIDO site._
+```{r}
+ecfu_MIDO_location <- here::here("results", "MIDO_ecfu_lm.png")
 
-```{r pressure, echo=FALSE}
-plot(pressure)
+include_graphics(ecfu_MIDO_location)
 ```
 
-Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
+</br>
+
+_Figure 3: Plots of biological score versus conductivity, E. coli cfu, nitrate concentration, and turbidity for the NORO site._
+```{r}
+bs_NORO_location <- here::here("results", "NORO_bs_lm.png")
+
+include_graphics(bs_NORO_location)
+```
+
+</br>
+
+_Figure 4: Plots of E. coli cfu versus conductivity, nitrate concentration, and turbidity for the NORO site._
+```{r}
+ecfu_NORO_location <- here::here("results", "NORO_ecfu_lm.png")
+
+include_graphics(ecfu_NORO_location)
+```
+
+</br>
+
+_Figure 5: Plots of biological score versus conductivity, E. coli cfu, nitrate concentration, and turbidity for the BICO site._
+```{r}
+bs_BICO_location <- here::here("results", "BICO_bs_lm.png")
+
+include_graphics(bs_BICO_location)
+```
+
+</br>
+
+_Figure 6: Plots of E. coli cfu versus conductivity, nitrate concentration, and turbidity for the BICO site._
+```{r}
+ecfu_BICO_location <- here::here("results", "BICO_ecfu_lm.png")
+
+include_graphics(ecfu_BICO_location)
+```

### 2021-11-15 code/analysis_code/analysis_uown.Rmd
Supplementary_Data - CC

@@
 date: "10/27/2021"
-output: html_document
+output:
+  word_document: default
+  html_document: default
 ---
@@
 _3. Rerun linear regressions for the different stream health indicator relationships and assess if more significant relationships exist._
-</br>
-#### Part 2: Linear Model Fit Analysis of MIDO Data
+
+</br>
+
+#### PART 2: Linear Model Fit Analysis of MIDO Data
 _1. Conduct test/train data splitting for MIDO Data._
@@
 
+</br>
+
+#### PART 3: LASSO Modeling MIDO Data
+_Conduct LASSO modeling of linear regression between Biological Score and all other variables and E.
+coli CFU and all other variables.
+Cross validation well be done by calculating RMSE for all models and compared to the RMSE for linear regression modeling of a Null model.
+Additionally, The best model will be selected using the R function select_best(), for which residuals will be calculated between predicted and actual data._
+
 ### Required packages
@@
 library(glmnet) #Model fitting
+library(knitr) #saving tables
 ```
@@
 
-ecfu_con_fit %>% 
+Ecoli_con_lm <- ecfu_con_fit %>% 
   extract_fit_parsnip() %>% 
@@
 
-ecfu_con_aug %>%
+Ecoli_con_RMSE <- ecfu_con_aug %>%
   rmse(truth = e.coli.cfu, .pred)
@@
 ```
-</br>
-
-#####Analysis here#####
 
@@
 
-ecfu_con_fit_test %>% 
+Ecoli_con_test_lm <- ecfu_con_fit_test %>% 
   extract_fit_parsnip() %>% 
@@
 ```
-#####Analysis here#####
 
@@
 
-ecfu_con_aug_test %>%
+Ecoli_con_RMSE_test <- ecfu_con_aug_test %>%
   rmse(truth = e.coli.cfu, .pred)
@@
 ```{r}
-#Create a location and save linear model for E. coli (cfu) in the MIDO train data.
-MIDO_ECFU_Con_fit_location <- here::here("results","MIDO_ECFU_Con_fit.rds")
-saveRDS(ecfu_con_fit, file = MIDO_ECFU_Con_fit_location)
-
-#Create a location and save linear model for E. coli (cfu) in the MIDO test data.
-MIDO_ECFU_Con_fit_test_location <- here::here("results","MIDO_ECFU_Con_fit_test.rds")
-saveRDS(ecfu_con_fit_test, file = MIDO_ECFU_Con_fit_test_location)
+#Create a table for E. coli (cfu) as a function of Conductivity Linear Fit Model in the MIDO train data.
+MIDO_ECFU_Con_fit <- kable(as.tibble(Ecoli_con_lm),
+      caption = "Linear Fit Model: E. Coli CFU predicted by Conductivity in the Trained MIDO Data Set")
+
+MIDO_ECFU_Con_fit
+
+#Create a table for E. coli (cfu) as a function of Conductivity Linear Fit Model in the MIDO test data.
+
+MIDO_ECFU_Con_fit_test <- kable(as.tibble(Ecoli_con_test_lm),
+      caption = "Linear Fit Model: E. Coli CFU predicted by Conductivity in the Test MIDO Data Set")
+
+MIDO_ECFU_Con_fit_test
 ```
@@
 ```{r}
-#Create a location and save RMSE for E. coli (cfu) in the MIDO train data.
-MIDO_ECFU_Con_RMSE_location <- here::here("results","MIDO_ECFU_Con_RMSE.rds")
-saveRDS(ecfu_con_aug, file = MIDO_ECFU_Con_RMSE_location)
-
-#Create a location and save linear RMSE for E. coli (cfu) in the MIDO test data.
-MIDO_ECFU_Con_RMSE_test_location <- here::here("results","MIDO_ECFU_Con_RMSE_test.rds")
-saveRDS(ecfu_con_aug_test, file = MIDO_ECFU_Con_RMSE_test_location)
+#Create a table for E. coli (cfu) as a function of Conductivity Linear Fit Model RMSE in the MIDO train data.
+MIDO_ECFU_Con_RMSE <- kable(as.tibble(Ecoli_con_RMSE),
+      caption = "Root Mean Square Error: E. Coli CFU predicted by Conductivity in the Trained MIDO Data Set")
+
+MIDO_ECFU_Con_RMSE
+
+#Create a table for E. coli (cfu) as a function of Conductivity Linear Fit Model RMSE in the MIDO test data.
+
+MIDO_ECFU_Con_RMSE_test <- kable(as.tibble(Ecoli_con_RMSE),
+      caption = "Root Mean Square Error: E. Coli CFU predicted by Conductivity in the Test MIDO Data Set")
+
+MIDO_ECFU_Con_RMSE_test
 ```
@@
 Null_Met <- collect_metrics(null_lm)
+
+Null_Met <- kable(Null_Met,
+      caption = "Root Mean Squared Error and Standard Deviation for Biological Score Null Linear Model")
 
@@
 #Showing best performing LASSO models
-lasso_resample %>%
+top_five_bs <- lasso_resample %>%
   show_best()
@@
   select_best()
+
+Best_LASSO_BS <- kable(top_five_bs,
+      caption = "Top 5 best performing LASSO Models for Biological Score as the Outcome")
+
+Best_LASSO_BS
 ```
@@
 lasso_pred_plot
-```
+
+#save plot
+lasso_pred_plot_BS_location <- here::here("results", "lasso_prediction_plot_BS.png")
+
+ggsave(lasso_pred_plot_BS_location)
+```
+
+</br>
+
 _plot residuals vs predictions_
@@
 
+Null_Met_EC <- kable(Null_Met_EC,
+      caption = "Root Mean Squared Error and Standard Deviation for E. coli cfu Null Linear Model")
+
 Null_Met_EC
@@
 #Showing best performing LASSO models
-lasso_resample_EC %>%
+top_five_ec <- lasso_resample_EC %>%
   show_best()
@@
   select_best()
+
+Best_LASSO_EC <- kable(top_five_ec,
+      caption = "Top 5 best performing LASSO Models for E. coli cfu as the Outcome")
+
+Best_LASSO_EC
 ```
@@
 lasso_pred_plot_EC
+
+#save plot
+lasso_pred_plot_EC_location <- here::here("results", "lasso_prediction_plot_EC.png")
+
+ggsave(lasso_pred_plot_EC_location)
 ```

### 2021-11-17 products/manuscript/Manuscript.Rmd
Miscellaneous_edits - CC
Intro & Supplementary File - CC

@@
 
+```{r, echo=FALSE, results='hide', message=FALSE}
+library(here)
+library(tidyverse)
+```
+
 # Summary/Abstract
+
+</br>
+
+_Stream health and water quality is typically measured by its physical characteristics (chemistry, surrounding land use area, channelization, etc.); Microbial (fecal indicator), and ecology (diversity of indicator macroinvertebrates).
+However, these categories are usually analyzed separately to assess stream health [cite].
+If the impact of urbanization on freshwater bodies is not monolithic, we could be spending a lot of time and effort on monitoring programs that are not effective.
+Studies have shown conflicting results on what variables like conductivity, turbidity, e.
+coli cfu (fecal coliform indicator), nutrient (nitrogen and phosphorus) concentration, and biodiversity indicator score actually say about water quality [cite].
+Therefore, a study was conducted assess how well all these parameters correlated with each other in the context of the Adopt-a-Stream method.
+Specifically, can you use one of the indicators mentioned (physical, microbial, and ecological) to describe the others to make a statement about overall stream health?_
+
+_Data was acquired from a publicly available spreadsheet of water quality measurements collected by the Upper Oconee Watershed Network (UOWN) between 2003 and 2020.
+Data originated from three distinct watersheds in the Athens- Clarke County area: Middle Oconee (MIDO), North Oconee (NORO), and Oconee (BIOC).
+The data were analyzed first by using bivariate linear modeling of the variables conductivity, turbidity, nitrate (no3) concentration, E.
+coli cfu, and biodiversity indicator score (biological score).
+E.
+coli cfu and biological score were used as outcomes to specifically see if easy to measure physical/chemical properties could predict biological variables to assess water quality.
+Second, the MIDO data were analyzed using LASSO linear modeling to construct the best combinations of variables to predict E.
+coli cfu and biological score._
+
+_Results indicated that the only significant bivariate relationships that existed were between E.
+coli cfu and conductivity for the MIDO site only.
+LASSO modeling produced models that did not fit the data any better than the null model that was calculated for cross validation.
+This indicates that physical/chemical water quality measurements cannot predict biological measurements.
+This suggests that the citizen science approach to water quality monitoring should be re-evaluated for sampling robustness or that the effect of urbanization is spatially and temporally different and may not captured well without more in-depth sampling.
+Future studies should investigate how specific aspects of urbanization actually impact our indicator variables of water quality and stream health._
 
@@
 
-_When monitoring freshwater streams, stream health is typically measured by its physical characteristics (chemistry, surrounding land use area, channelization, etc.); Microbial (fecal indicator), and ecology (diversity of indicator macroinvertebrates).
-However, these categories are usually analyzed separately to assess stream health.
-As such, my question is how well are these parameters correlated with each other.
-Specifically, can you use one of the indicators mentioned (physical, microbial, and ecological) to describe both of the others to make a statement about overall stream health?_
+_The Earth's surface contains a relative proportion of approximately 3% freshwater to 97% salt water.
+As such, that three percent is supposedly meant to sustain a growing global population of 9 billion people.
+In a world where water scarcity is more and more common, our freshwater reserves are also threatened by pollution by urbanization.
+Urbanization surrounding a freshwater body creates a watershed dominated by impervious surface in the form of concrete and asphalt [cite].
+Studies have shown that there is a significant negative relationship between impervious surface coverage and water quality.
+This is most likely due to accumulation and concentration of run-off pollutants from an entire watershed into a single drainage basin (i.e.
+freshwater body)[cite].
+Because the urban world continues to grow with population, it is likely that we will see a proportional decrease in water quality in our freshwater bodies as a result.
+Therefore, a rapid and consistent monitoring system needs to be in place to effectively allocate resources to remediating body of water in the midst of being polluted._
 
-</br>
+_Currently, stream health and water quality is typically measured by its physical characteristics (chemistry, surrounding land use area, channelization, etc.); Microbial (fecal indicator), and ecology (diversity of indicator macroinvertebrates).
+Typically, these measurements include conductivity (measure of dissolved ions), dissolved oxygen, temperature, pH, nitrate and phosphorus concentration, turbidity (measure of suspended particulate matter in the water column), fecal indicator bacteria (measure of waste pollution), and benthic macroinvertebrate diversity indices (measure of impervious surface run-off on sensitive aquatic taxa) [cite].
+However, these categories are usually analyzed separately to assess stream health [cite].
+As such, it is hard to evaluate whether urbanization affects watersheds and water quality in fairly similar ways, regardless of location.
+Evaluating how uniform the impact of urbanization is on different watersheds would allow for a cookie-cutter monitoring and remediation method, which would allow implementation to be fairly easy for environmental managers.
+On the other hand, if the impact of urbanization on freshwater bodies is not monolithic, we could be spending a lot of time and effort on monitoring programs that are not effective.
+Currently, citizen science groups, like Adopt-a-Stream in Georgia, are collecting data on many watersheds throughout the world in an effort to consistently monitor water quality in a community.
+However, studies have shown conflicting results on what variables like conductivity, turbidity, e.
+coli cfu (fecal coliform indicator), nutrient (nitrogen and phosphorus) concentration, and biodiversity indicator score actually say about water quality [cite].
+Therefore, my question is how well are these parameters correlated with each other in the context of the Adopt-a-Stream method.
+Specifically, can you use one of the indicators mentioned (physical, microbial, and ecological) to describe the others to make a statement about overall stream health?_
 
-_I would expect that because the all three are correlated with urbanization/anthropocentric impact, a negative impact in one indicator category will indicate a correlated negative impact in the other categories.
+_I would expect that because the all three are correlated with urbanization/anthropocentric impact, a negative impact in one indicator category will indicate a correlated negative impact to stream health in the other categories.
 The only caveat is that because level of noise in stream health data, significant correlations may be difficult to find._
@@
 
-</br>
-
 _2. Designate NA's as actually missing values instead of factor data._
-
-</br>
 
@@
 
-</br>
-
-_There is 3,075 rows and 12 columns...
-So, a lot of data.
+_There is 3,075 rows and 12 columns.
 This is because data has been collected for ~20 years.
@@
 
-</br>
-
-_4. Use NA assessment to see determine which variables to remove from assessment. This was done using the is_na function._
-_Several of the columns have more than half of the data missing.
+_4.
+Use NA assessment to see determine which variables to remove from assessment.
+This was done using the is_na function.
+Several of the columns have more than half of the data missing.
 This is pretty common for citizen science data, seeing as collection is opportunistic.
@@
 I removed NA data using the na.omit() function_
-
-</br>
 
@@
 
- </br>
-
 _6. Separate the data into three separate dataframes based on location ID: "MIDO", "NORO", and "BICO". This is to separate out any bias based on location of the streams within ACC._
-
-</br>
 

### 2021-11-17 results/Supplementary_Data.Rmd
Intro & Supplementary File - CC

@@
 ```
+
+</br>
+
+_Figure 7: Scatter plot of LASSO model predicted Biological Score data as a function of actual Biological Score from the MIDO data set._
+```{r}
+lasso_prediction_plot_BS_location <- here::here("results", "lasso_prediction_plot_BS.png")
+
+include_graphics(lasso_prediction_plot_BS_location)
+```
+</br>
+
+_Figure 8: Scatter plot of LASSO model predicted E. coli cfu data as a function of actual E. coli cfu from the MIDO data set._
+```{r}
+lasso_prediction_plot_EC_location <- here::here("results", "lasso_prediction_plot_EC.png")
+
+include_graphics(lasso_prediction_plot_EC_location)
+```
+
+</br>
+
+```{r}
+MIDO_ecfu_con_fit_location <- here::here("results", "MIDO_ecfu_con_fit.png")
+
+include_graphics(MIDO_ecfu_con_fit_location)
+```
+</br>
+_Table 1_
+
+</br>
+
+```{r}
+MIDO_ecfu_con_RMSE_location <- here::here("results", "MIDO_ecfu_con_RMSE.png")
+
+include_graphics(MIDO_ecfu_con_RMSE_location)
+```
+</br>
+_Table 2_
+
+</br>
+
+```{r}
+Null_bs_RMSE_location <- here::here("results", "Null_bs_RMSE.png")
+
+include_graphics(Null_bs_RMSE_location)
+```
+</br>
+_Table 3_
+
+</br>
+
+```{r}
+Null_ecfu_RMSE_location <- here::here("results", "Null_ecfu_RMSE.png")
+
+include_graphics(Null_ecfu_RMSE_location)
+```
+</br>
+_Table 4_
+
+</br>
+
+```{r}
+MIDO_bs_LASSO_top5_location <- here::here("results", "MIDO_bs_LASSO_top5.png")
+
+include_graphics(MIDO_bs_LASSO_top5_location)
+```
+</br>
+_Table 5_
+
+</br>
+
+```{r}
+MIDO_ecfu_LASSO_top5_location <- here::here("results", "MIDO_ecfu_LASSO_top5.png")
+
+include_graphics(MIDO_ecfu_LASSO_top5_location)
+```
+</br>
+_Table 6_

### 2021-11-18 code/analysis_code/analysisscript.R
miscellaneous_edits

@@ -1,8 +1,0 @@
-###############################
-# analysis script
-#
-#this script loads the processed, cleaned data, does complex analysis
-#for Carter Coleman's UOWN indicator of Stream Health Project.
-
-
-  

### 2021-11-29 code/analysis_code/analysis_uown.Rmd
Include References & Formatting - CC

@@
 ```{r}
-library(tidyverse) #Working with multiple Tidy packages
-library(tidymodels) #Building models
+library(tidyverse) #working with multiple Tidy packages
+library(tidymodels) #building models
 library(dplyr) #data manipulation
 library(here) #setting pathways for saving files
-library(rpart) #Model fitting
-library(ranger) #Model fitting
-library(glmnet) #Model fitting
+library(rpart) #model fitting
+library(ranger) #model fitting
+library(glmnet) #model fitting
 library(knitr) #saving tables
+library(gridExtra) #grouping figure outputs together
 ```
@@
 _Outcome: Biological Score_
-```{r}
+```{r, fig.width=10}
+#Set save file path parameters
+png(filename = here::here("results", "MIDO_BS_OUT.png"), width = 10, height = 6, units = "in", res = 300)
+
+par(mfrow=c(2,2)) #group plots together using base r
+
 #Creates a dataframe that stores the cook's distance analysis of influential data points.
 bs_con_cd <- cooks.distance(bs_con_lm)
-plot(bs_con_cd, pch="*", cex=2, main="MIDO: Conductivity vs. Biological Score with Outliers")
+bs_con_cd_plot <- plot(bs_con_cd, pch="*", cex=2, main="MIDO: Conductivity vs. Biological Score with Outliers")
 abline(h = 4/sample_size_MIDO, col="red")
@@
 bs_ecfu_cd <- cooks.distance(bs_ecfu_lm)
-plot(bs_ecfu_cd, pch="*", cex=2, main="MIDO: E. coli CFU vs. Biological Score with Outliers")
+bs_ecfu_cd_plot <- plot(bs_ecfu_cd, pch="*", cex=2, main="MIDO: E. coli CFU vs. Biological Score with Outliers")
 abline(h = 4/sample_size_MIDO, col="red")
@@
 bs_no3_cd <- cooks.distance(bs_no3_lm)
-plot(bs_no3_cd, pch="*", cex=2, main="MIDO: NO3 vs. Biological Score with Outliers")
+bs_no3_cd_plot <- plot(bs_no3_cd, pch="*", cex=2, main="MIDO: NO3 vs. Biological Score with Outliers")
 abline(h = 4/sample_size_MIDO, col="red")
@@
 bs_turb_cd <- cooks.distance(bs_turb_lm)
-plot(bs_turb_cd, pch="*", cex=2, main="MIDO: Turbidity vs. Biological Score with Outliers")
+bs_turb_cd_plot <- plot(bs_turb_cd, pch="*", cex=2, main="MIDO: Turbidity vs. Biological Score with Outliers")
 abline(h = 4/sample_size_MIDO, col="red")
+
+dev.off()
 ```
@@
 _Outcome: E.coli CFU_
-```{r}
+```{r, fig.width=10}
+#Set save file path parameters
+png(filename = here::here("results", "MIDO_EC_OUT.png"), width = 10, height = 6, units = "in", res = 300)
+
+par(mfrow=c(2,2)) #group plots together using base r
+
 #Creates a dataframe that stores the cook's distance analysis of influential data points.
@@
 abline(h = 4/sample_size_MIDO, col="red")
+
+dev.off()
 ```
@@
 _Outcome: Biological Score_
-```{r}
+```{r, fig.width=10}
+#Set save file path parameters
+png(filename = here::here("results", "NORO_BS_OUT.png"), width = 10, height = 6, units = "in", res = 300)
+
+par(mfrow=c(2,2)) #group plots together using base r
+
 #Creates a dataframe that stores the cook's distance analysis of influential data points.
@@
 abline(h = 4/sample_size_NORO, col="red")
+
+dev.off()
 ```
@@
 _Outcome: E.coli CFU_
-```{r}
+```{r, fig.width=10}
+#Set save file path parameters
+png(filename = here::here("results", "NORO_EC_OUT.png"), width = 10, height = 6, units = "in", res = 300)
+
+par(mfrow=c(2,2)) #group plots together using base r
+
 #Creates a dataframe that stores the cook's distance analysis of influential data points.
@@
 _Outcome: Biological Score_
-```{r}
+```{r, fig.width=10}
+#Set save file path parameters
+png(filename = here::here("results", "BICO_BS_OUT.png"), width = 10, height = 6, units = "in", res = 300)
+
+par(mfrow=c(2,2)) #group plots together using base r
+
 #Creates a dataframe that stores the cook's distance analysis of influential data points.
@@
 abline(h = 4/sample_size_BICO, col="red")
+
+dev.off()
 ```
@@
 _Outcome: E.coli CFU_
-```{r}
+```{r, fig.width=10}
+#Set save file path parameters
+png(filename = here::here("results", "BICO_EC_OUT.png"), width = 10, height = 6, units = "in", res = 300)
+
+par(mfrow=c(2,2)) #group plots together using base r
+
 #Creates a dataframe that stores the cook's distance analysis of influential data points.
@@
 ```{r}
+
 #Create a table for E. coli (cfu) as a function of Conductivity Linear Fit Model in the MIDO train data.
 MIDO_ECFU_Con_fit <- kable(as.tibble(Ecoli_con_lm),
-      caption = "Linear Fit Model: E. Coli CFU predicted by Conductivity in the Trained MIDO Data Set")
-
-MIDO_ECFU_Con_fit
+      caption = "Table 1: Linear Fit Model: E. Coli CFU predicted by Conductivity in the Trained MIDO Data Set")
+
+MIDO_ECFU_Con_location <- here::here("results", "MIDO_ECFU_Con_location.RDS")
+
+saveRDS(MIDO_ECFU_Con_fit, MIDO_ECFU_Con_location)
 
@@
 MIDO_ECFU_Con_RMSE <- kable(as.tibble(Ecoli_con_RMSE),
-      caption = "Root Mean Square Error: E. Coli CFU predicted by Conductivity in the Trained MIDO Data Set")
-
-MIDO_ECFU_Con_RMSE
+      caption = "Table 2: Root Mean Square Error: E. Coli CFU predicted by Conductivity in the Trained MIDO Data Set")
+
+MIDO_ECFU_Con_RMSE_location <- here::here("results", "MIDO_ECFU_Con_RMSE_location.RDS")
+
+saveRDS(MIDO_ECFU_Con_RMSE, MIDO_ECFU_Con_RMSE_location)
 
@@
 Null_Met <- kable(Null_Met,
-      caption = "Root Mean Squared Error and Standard Deviation for Biological Score Null Linear Model")
-
-Null_Met
+      caption = "Table 3: Root Mean Squared Error and Standard Deviation for Biological Score Null Linear Model")
+
+BS_Null_Mod_Location <- here::here("results", "BS_Null_Mod_Location.RDS")
+
+saveRDS(Null_Met, BS_Null_Mod_Location)
 ```
@@
 Best_LASSO_BS <- kable(top_five_bs,
-      caption = "Top 5 best performing LASSO Models for Biological Score as the Outcome")
-
-Best_LASSO_BS
+      caption = "Table 4: Top 5 best performing LASSO Models for Biological Score as the Outcome")
+
+Best_LASSO_BS_location <- here::here("results", "Best_LASSO_BS_location.RDS")
+
+saveRDS(Best_LASSO_BS, Best_LASSO_BS_location)
 ```
@@
 Null_Met_EC <- kable(Null_Met_EC,
-      caption = "Root Mean Squared Error and Standard Deviation for E. coli cfu Null Linear Model")
-
-Null_Met_EC
+      caption = "Table 5: Root Mean Squared Error and Standard Deviation for E. coli cfu Null Linear Model")
+
+Null_Met_EC_location <- here::here("results", "Null_Met_EC_location.RDS")
+
+saveRDS(Null_Met_EC, Null_Met_EC_location)
 ```
@@
 Best_LASSO_EC <- kable(top_five_ec,
-      caption = "Top 5 best performing LASSO Models for E. coli cfu as the Outcome")
-
-Best_LASSO_EC
+      caption = "Table 6: Top 5 best performing LASSO Models for E. coli cfu as the Outcome")
+
+Best_LASSO_EC_location <- here::here("results", "Best_LASSO_EC_location.RDS")
+
+saveRDS(Best_LASSO_EC, Best_LASSO_EC_location)
 ```

### 2021-11-29 code/processing_code/Data_Exploration.Rmd
Include References & Formatting - CC

@@
 ```{r}
-library(dplyr) 
-library(here)
-library(tidyverse)
-library(tidyr)
-library(readr)
-library(stringr)
-library(gridExtra)
+library(dplyr) #data manipulation
+library(here) #setting pathways for saving files
+library(tidyverse) #Working with multiple Tidy packages
+library(tidyr) #data cleaning
+library(readr) #read in files
+library(stringr) #working with strings
+library(gridExtra) #grouping figure outputs together
 ```

### 2021-11-29 products/manuscript/Manuscript.Rmd
Include References & Formatting - CC

@@
   bookdown::word_document2:
+    reference_docx: ../style_template_final.docx 
     toc: false
@@
     toc: false
-bibliography: ../references.bib
-csl: ../apa.csl
+bibliography: ../References (2).bib
+csl: ../elsevier_style.csl
 ---
@@
 ```{r, echo=FALSE, results='hide', message=FALSE}
-library(here)
-library(tidyverse)
+library(knitr)    # For knitting document and include_graphics function
+library(ggplot2)  # For plotting
+library(png)      # For grabbing the dimensions of png files
 ```
 
-# Summary/Abstract
+# Abstract
 
@@
 
-_Stream health and water quality is typically measured by its physical characteristics (chemistry, surrounding land use area, channelization, etc.); Microbial (fecal indicator), and ecology (diversity of indicator macroinvertebrates).
-However, these categories are usually analyzed separately to assess stream health [cite].
+Stream health and water quality is typically measured by its physical characteristics (chemistry, surrounding land use area, channelization, etc.); Microbial (fecal indicator), and ecology (diversity of indicator macroinvertebrates).
+However, these categories are usually analyzed separately to assess stream health.
 If the impact of urbanization on freshwater bodies is not monolithic, we could be spending a lot of time and effort on monitoring programs that are not effective.
 Studies have shown conflicting results on what variables like conductivity, turbidity, e.
-coli cfu (fecal coliform indicator), nutrient (nitrogen and phosphorus) concentration, and biodiversity indicator score actually say about water quality [cite].
+coli cfu (fecal coliform indicator), nutrient (nitrogen and phosphorus) concentration, and biodiversity indicator score actually say about water quality.
 Therefore, a study was conducted assess how well all these parameters correlated with each other in the context of the Adopt-a-Stream method.
-Specifically, can you use one of the indicators mentioned (physical, microbial, and ecological) to describe the others to make a statement about overall stream health?_
+Specifically, can you use one of the indicators mentioned (physical, microbial, and ecological) to describe the others to make a statement about overall stream health?
 
-_Data was acquired from a publicly available spreadsheet of water quality measurements collected by the Upper Oconee Watershed Network (UOWN) between 2003 and 2020.
+Data was acquired from a publicly available spreadsheet of water quality measurements collected by the Upper Oconee Watershed Network (UOWN) between 2003 and 2020.
 Data originated from three distinct watersheds in the Athens- Clarke County area: Middle Oconee (MIDO), North Oconee (NORO), and Oconee (BIOC).
@@
 Second, the MIDO data were analyzed using LASSO linear modeling to construct the best combinations of variables to predict E.
-coli cfu and biological score._
+coli cfu and biological score.
 
-_Results indicated that the only significant bivariate relationships that existed were between E.
+Results indicated that the only significant bivariate relationships that existed were between E.
 coli cfu and conductivity for the MIDO site only.
@@
 This suggests that the citizen science approach to water quality monitoring should be re-evaluated for sampling robustness or that the effect of urbanization is spatially and temporally different and may not captured well without more in-depth sampling.
-Future studies should investigate how specific aspects of urbanization actually impact our indicator variables of water quality and stream health._
+Future studies should investigate how specific aspects of urbanization actually impact our indicator variables of water quality and stream health.
 
@@
 
-_The Earth's surface contains a relative proportion of approximately 3% freshwater to 97% salt water.
+The Earth's surface contains a relative proportion of approximately 3% freshwater to 97% salt water.
 As such, that three percent is supposedly meant to sustain a growing global population of 9 billion people.
-In a world where water scarcity is more and more common, our freshwater reserves are also threatened by pollution by urbanization.
-Urbanization surrounding a freshwater body creates a watershed dominated by impervious surface in the form of concrete and asphalt [cite].
-Studies have shown that there is a significant negative relationship between impervious surface coverage and water quality.
+In a world where water scarcity is more and more common due to increasingly warmer climate (@Winter2017), our freshwater reserves are also threatened by pollution by urbanization.
+Urbanization surrounding a freshwater body creates a watershed dominated by impervious surface in the form of concrete and asphalt (@Bouizrou2021).
+Studies have shown that there is a significant negative relationship between impervious surface coverage and water quality (@Birawat2021).
 This is most likely due to accumulation and concentration of run-off pollutants from an entire watershed into a single drainage basin (i.e.
-freshwater body)[cite].
-Because the urban world continues to grow with population, it is likely that we will see a proportional decrease in water quality in our freshwater bodies as a result.
-Therefore, a rapid and consistent monitoring system needs to be in place to effectively allocate resources to remediating body of water in the midst of being polluted._
+freshwater body)(@Klein1979; @Wolfram2021).
+Because the urban world continues to grow with population, it is likely that we will see a proportional decrease in water quality in our freshwater bodies as a result (@Wolfram2021).
+Therefore, a rapid and consistent monitoring system needs to be in place to effectively allocate resources to remediating body of water in the midst of being polluted.
 
-_Currently, stream health and water quality is typically measured by its physical characteristics (chemistry, surrounding land use area, channelization, etc.); Microbial (fecal indicator), and ecology (diversity of indicator macroinvertebrates).
-Typically, these measurements include conductivity (measure of dissolved ions), dissolved oxygen, temperature, pH, nitrate and phosphorus concentration, turbidity (measure of suspended particulate matter in the water column), fecal indicator bacteria (measure of waste pollution), and benthic macroinvertebrate diversity indices (measure of impervious surface run-off on sensitive aquatic taxa) [cite].
-However, these categories are usually analyzed separately to assess stream health [cite].
-As such, it is hard to evaluate whether urbanization affects watersheds and water quality in fairly similar ways, regardless of location.
+Currently, stream health and water quality are typically measured by their physical characteristics (chemistry, surrounding land use area, channelization, etc.); Microbial (fecal indicator), and ecology (diversity of indicator macroinvertebrates) (@Clapcott2012; @Flinders2008; @Alvareda2020).
+Typically, these measurements include conductivity (measure of dissolved ions), dissolved oxygen, temperature, pH, nitrate and phosphorus concentration, turbidity (measure of suspended particulate matter in the water column), fecal indicator bacteria (measure of waste pollution), and benthic macroinvertebrate diversity indices (measure of impervious surface run-off on sensitive aquatic taxa) (@Clapcott2012; @Flinders2008; @Alvareda2020).
+While Biological Indicatior Score is used to measure the combination of stream chemical and physical properties (@Flinders2008), it is hard to evaluate whether urbanization affects watersheds and water quality uniformly (@Wu2020).
 Evaluating how uniform the impact of urbanization is on different watersheds would allow for a cookie-cutter monitoring and remediation method, which would allow implementation to be fairly easy for environmental managers.
@@
 However, studies have shown conflicting results on what variables like conductivity, turbidity, e.
-coli cfu (fecal coliform indicator), nutrient (nitrogen and phosphorus) concentration, and biodiversity indicator score actually say about water quality [cite].
-Therefore, my question is how well are these parameters correlated with each other in the context of the Adopt-a-Stream method.
-Specifically, can you use one of the indicators mentioned (physical, microbial, and ecological) to describe the others to make a statement about overall stream health?_
+coli cfu (fecal coliform indicator), nutrient (nitrogen and phosphorus) concentration, and biodiversity indicator score actually say about water quality (@Boulton1999).
+Therefore, I aim to answer how well are these parameters correlated with each other in the context of the Adopt-a-Stream method.
+Specifically, can you use one of the indicators mentioned (physical, microbial, and ecological) to describe the others to make a statement about overall stream health?
 
-_I would expect that because the all three are correlated with urbanization/anthropocentric impact, a negative impact in one indicator category will indicate a correlated negative impact to stream health in the other categories.
-The only caveat is that because level of noise in stream health data, significant correlations may be difficult to find._
+I would expect that because the all three are correlated with urbanization/anthropocentric impact, a negative impact in one indicator category will indicate a correlated negative impact to stream health in the other categories.
+The only caveat is that because level of noise in stream health data, significant correlations may be difficult to find.
 
@@
 
-# Methods and Results
+# Methods
 
-_UOWN has provided a great deal of data dating back to 2003.
+UOWN has provided a great deal of data dating back to 2003.
 First,I will need to wrangle missing data values that are a result of inconsistent sampling efforts.
@@
 Next, since biological scores are used as a measure of broader ecological health, I will run t-tests on if the other factors significantly effect biological score.
-Finally, I will test the interaction of all three via multivariate analysis._
+Finally, I will test the interaction of all three via multivariate analysis.
 
@@
 
-_The data I am using is the publically avaialable data provided by the Upper Oconee Watershed Network (UOWN).
+The data I am using is the publically avaialable data provided by the Upper Oconee Watershed Network (UOWN).
 This data is part of a citizen science training effort for routine watershed-scale monitoring of the upper oconee river watershed.
@@
 The raw data shows 3,075 observations before cleaning.
-However, while cleaning I had to eliminate "po4.mgl" and "temperature" due to significant amounts of missing data._
+However, while cleaning I had to eliminate "po4.mgl" and "temperature" due to significant amounts of missing data.
 
@@
 
-_Below is the cleaned data produced from our data processing. Data was processed using the processingscript.R file and includes the following steps:_
-_1. Load rawdata from excel file. Note the use of the here() package and not absolute paths_
+Below is the cleaned data produced from our data processing. Data was processed using the processingscript.R file and includes the following steps:
 
-_2. Designate NA's as actually missing values instead of factor data._
+1. Load rawdata from excel file. Note the use of the here() package and not absolute paths.
 
-_3.
+2. Designate NA's as actually missing values instead of factor data.
+
+3.
 Select variables of interest.
@@
 Such variables are listed here: WSID, biological_score, conductivity.uscm, turbidity.ntu, po4.mgL, no3.mgL, pH, temperature.c,	e.coli.cfu, month, year,	and day.
-This was done using the Select() function._
+This was done using the Select() function.
 
-_There is 3,075 rows and 12 columns.
+There is 3,075 rows and 12 columns.
 This is because data has been collected for ~20 years.
 Unfortunately, I am seeing a lot of NA's in the data.
-Because this is citizen science monitoring data, NA most likely means the data wasn't collected during that event and is a form of human error bias._
+Because this is citizen science monitoring data, NA most likely means the data wasn't collected during that event and is a form of human error bias.
 
-_4.
+4.
 Use NA assessment to see determine which variables to remove from assessment.
@@
 Finally, I decided to remove all missing values and see the condition of the data knowing that there are now no missing values.
-I removed NA data using the na.omit() function_
+I removed NA data using the na.omit() function.
 
-_5.
+5.
 Change all numeric columns to be read as numeric instead of character.
 This sometimes happens when R tries to read in data from an external source, but is an easy fix with the as.numeric function.
-This is important because my analysis uses linear regression functions and they are coded to only deal with numeric values._
+This is important because my analysis uses linear regression functions and they are coded to only deal with numeric values.
 
-_6. Separate the data into three separate dataframes based on location ID: "MIDO", "NORO", and "BICO". This is to separate out any bias based on location of the streams within ACC._
+6. Separate the data into three separate data frames based on location ID: "MIDO", "NORO", and "BICO". This is to separate out any bias based on location of the streams within ACC.
 
-_7.
+7.
 Because the number of rows between each site data frame and between years is not consistent, measuring between sites and through time may be biased.
 That being said, my original question refers to the correlation between different indicators of stream health, not how they change spatial or temporally.
-Therefore, these space and time inconsistencies won't affect the analyses._
-
-</br>
+Therefore, these space and time inconsistencies won't affect the analyses.
 
@@
 
-_The following is an exploration of Stream Health Indicator correlation and comparison. The exploration will be structured as follows:_
-_Each dataframe (MIDO, NORO, and BICO) will be analyzed separately.
+The following is an exploration of Stream Health Indicator correlation and comparison. The exploration will be structured as follows:
+
+Each dataframe (MIDO, NORO, and BICO) will be analyzed separately.
 Within each dataframe, there are data explaining biological indicator values, stream chemistry, and fecal indicators over time.
@@
 I plan to do a series of correlative analysis showing if there is some sort of relationship between the three indicator types to help inform further analysis via principle component, multivariate analysis, etc.
-Therefore, the goal of this exploration is basically to summarize how these indicators correlate._
+Therefore, the goal of this exploration is basically to summarize how these indicators correlate.
 
@@
 
-_At this point,the main take-aways from data analysis thus far is that correlation of stream health indicators showed that most interactions were non-significant, with the exception of E.
+At this point, the main take-aways from data analysis thus far is that correlation of stream health indicators showed that most interactions were non-significant, with the exception of E.
 coli (cfu) and turbidity.
-Additionally, the data set is most likely not skewed by influential outliers._
+Additionally, the data set is most likely not skewed by influential outliers.
 
-</br>
-
-_Following correlation analysis during exploration, focus was shifted to just the MIDO dataset.
+Following correlation analysis during exploration, focus was shifted to just the MIDO dataset.
 As you can tell, working with many variables in three different data sets can be quite cumbersome to read through.
@@
 For comparison, MIDO has an n = 88; NORO has an n = 35; and BICO has an n = 23.
-The only thing we are missing by dropping the other sites is assessing site-by-site comparison, which I am choosing not to focus on anyway._
+The only thing we are missing by dropping the other sites is assessing site-by-site comparison, which I am choosing not to focus on anyway.
 
-</br>
-
-_Finally, LASSO analyses were performed with either biological score or E.
+Finally, LASSO analyses were performed with either biological score or E.
 coli cfu as the outcome of choice.
 The LASSO model analyzed which combination of stream health descriptive variables are the best for predicting either biological score or E.
-coli cfu values for the MIDO site through time._
+coli cfu values for the MIDO site through time.
 
-</br>
+# Results
 
-_The results indicated that only the relationship between E.
+Results from exploratory data analysis indicates that most interactions between stream health indicators were non-significant, with the exception of E.
+coli (cfu) and turbidity.
+Additionally, the data set is most likely not skewed by influential outliers.
+
+In addition to non-significant interactions, the relationships (slopes) between either biological score or E. coli cfu and stream physical and chemical properties varied inconsistently between sites.
+
+Focusing on just biological score as the outcome variable biological score (supplementary figures 1-3), the interaction with conductivity was determined to be negative for the MIDO and NORO site.
+However, the BICO site had a positive interaction.
+Next, the regression plotting found a negative interaction with E.
+coli cfu for the MIDO and BICO sites.
+But, there was a positive interaction for the NORO site.
+For nitrate concentration, the interaction was negative for all sites.
+Finally, the interaction with turbidity was negative for the BICO site, while a positive interaction was found for the MIDO and NORO sites.
+
+Focusing on E.
+coli cfu as the outcome of interest (figures 1-3 below), the interaction with conductivity was negative for MIDO and NORO.
+However, the BICO site had a positive interaction.
+The interaction with nitrate concentration was negative for all three sites.
+Finally the interaction with turbidity was positive at all three sites.
+
+_Figure 1: Plots of E. coli cfu versus conductivity, nitrate concentration, and turbidity for the MIDO site._
+```{r, echo=FALSE, message=FALSE}
+ecfu_MIDO_location <- here::here("results", "MIDO_ecfu_lm.png")
+
+include_graphics(ecfu_MIDO_location)
+```
+
+_Figure 2: Plots of E. coli cfu versus conductivity, nitrate concentration, and turbidity for the NORO site._
+```{r, echo=FALSE, message=FALSE}
+ecfu_NORO_location <- here::here("results", "NORO_ecfu_lm.png")
+
+include_graphics(ecfu_NORO_location)
+```
+
+_Figure 3: Plots of E. coli cfu versus conductivity, nitrate concentration, and turbidity for the BICO site._
+```{r, echo=FALSE, message=FALSE}
+ecfu_BICO_location <- here::here("results", "BICO_ecfu_lm.png")
+
+include_graphics(ecfu_BICO_location)
+```
+
+The results indicated that only the relationship between E.
 Coli (cfu) and Conductivity were significant for both the train data (p = 1.374747e-05) and test data (p = 1.89214e-06).
@@
 Finally, all RMSE values (actual and predicted data models) were compared to a null model for both biological score and E.
-coli cfu to evaluate model performance._
+coli cfu to evaluate model performance.
+
+```{r, echo=FALSE, message=FALSE}
+MIDO_ecfu_con_fit_location <- here::here("results", "MIDO_ECFU_Con_location.RDS")
+
+MIDO_ecfu_con_fit <- readRDS(MIDO_ecfu_con_fit_location)
+
+MIDO_ecfu_con_fit
+```
 
@@
 
-_The null linear models for biological score and E.
+```{r, echo=FALSE, message=FALSE}
+MIDO_ECFU_Con_RMSE_location <- here::here("results", "MIDO_ECFU_Con_RMSE_location.RDS")
+
+MIDO_ECFU_Con_RMSE <- readRDS(MIDO_ECFU_Con_RMSE_location)
+
+MIDO_ECFU_Con_RMSE
+```
+
+</br>
+
+The null linear models for biological score and E.
 coli cfu produced a RMSE = 6.63, with a standard deviation of 0.13 and RMSE = 0.52, with a standard deviation of 0.014, respectively.
@@
 However, based on comparative RMSE scores, model 41 does not really perform any better than the null model.
-This indicates that it is a bad fit to the MIDO data._
+This indicates that it is a bad fit to the MIDO data.
+
+```{r, echo=FALSE, message=FALSE}
+BS_Null_Mod_Location <- here::here("results", "BS_Null_Mod_Location.RDS")
+
+BS_Null_Mod <- readRDS(BS_Null_Mod_Location)
+
+BS_Null_Mod
+```
 
@@
 
-# Discussion
+```{r, echo=FALSE, message=FALSE}
+Best_LASSO_BS_location <- here::here("results", "Best_LASSO_BS_location.RDS")
+
+Best_LASSO_BS <- readRDS(Best_LASSO_BS_location)
+
+Best_LASSO_BS
+```
 
@@
 
-_Based on the linear modeling exploratory analysis and LASSO modeling for best combination of predictors, the results indicate that combinations of the variables biological score, E.
+```{r, echo=FALSE, message=FALSE}
+Null_Met_EC_location <- here::here("results", "Null_Met_EC_location.RDS")
+
+Null_Met_EC <- readRDS(Null_Met_EC_location)
+
+Null_Met_EC
+```
+
+</br>
+
+```{r, echo=FALSE, message=FALSE}
+Best_LASSO_EC_location <- here::here("results", "Best_LASSO_EC_location.RDS")
+
+Best_LASSO_EC <- readRDS(Best_LASSO_EC_location)
+
+Best_LASSO_EC
+```
+
+</br>
+
+_Figure 4: Scatter plot of LASSO model predicted Biological Score data as a function of actual Biological Score from the MIDO data set._
+```{r, echo=FALSE, message=FALSE}
+lasso_prediction_plot_BS_location <- here::here("results", "lasso_prediction_plot_BS.png")
+
+include_graphics(lasso_prediction_plot_BS_location)
+```
+
+_Figure 5: Scatter plot of LASSO model predicted E. coli cfu data as a function of actual E. coli cfu from the MIDO data set._
+```{r, echo=FALSE, message=FALSE}
+lasso_prediction_plot_EC_location <- here::here("results", "lasso_prediction_plot_EC.png")
+
+include_graphics(lasso_prediction_plot_EC_location)
+```
+
+# Discussion
+
+Based on the inconsistent linear modeling exploratory analysis and LASSO modeling for best combination of predictors, the results indicate that combinations of the variables biological score, E.
 coli cfu, nitrate concentration, conductivity, and turbidity are not able to predict values for biological score or E.
 coli cfu.
-This indicates that one cannot predict one measure of stream health from another.
+This indicates that one cannot predict either Biological Score or E.
+coli cfu from any combination of stream physical and chemical properties.
 This is seen by a lack of consistent significant correlations between each stream health predictor and either biological score or E.
@@
 While confidence in values may be greater in this case, there is know a shortage of labor due to need for a specialized skill set to process and quantify E.
-coli cfu values._
-
-</br>
+coli cfu values.
 
@@
 
-</br>
-
-_Based the results of this study, I would recommend that existing methodology for UOWN stream health characterization and monitoring be re-evaluated.
+Based the results of this study, I would recommend that existing methodology for UOWN stream health characterization and monitoring be re-evaluated.
 The issue being that there seems to be a disparity in sampling effort, both in quality of samples and time spent sampling, for each different stream health metric.
@@
 coli cfu so citizen based monitoring does not have to collect these data while still having easier to measure variables that can be used to describe stream health in terms of biological score and E.
-coli cfu._
+coli cfu.
 

### 2021-11-29 products/manuscript/Supplementary_Data.Rmd
Include References & Formatting - CC

---
title: "Supplementary Data"
author: "C. Coleman"
date: "11/12/2021"
output:
  word_document: default
  html_document: default
---

#### The following markdown serves as the supplementary materials for the UOWN Stream Health Indicators Project. Included materials are all figures relevant to my manuscript and were derived from the analysis code.

```{r,echo=FALSE, results='hide', message=FALSE}
library(knitr)    # For knitting document and include_graphics function
library(ggplot2)  # For plotting
library(png)      # For grabbing the dimensions of png files
```

_Supplementary Figure 1: Plots of biological score versus conductivity, E. coli cfu, nitrate concentration, and turbidity for the MIDO site._
```{r, echo=FALSE, message=FALSE}
bs_MIDO_location <- here::here("results", "MIDO_bs_lm.png")

include_graphics(bs_MIDO_location)
```

_Supplementary Figure 2: Plots of biological score versus conductivity, E. coli cfu, nitrate concentration, and turbidity for the NORO site._
```{r, echo=FALSE, message=FALSE}
bs_NORO_location <- here::here("results", "NORO_bs_lm.png")

include_graphics(bs_NORO_location)
```

_Supplementary Figure 3: Plots of biological score versus conductivity, E. coli cfu, nitrate concentration, and turbidity for the BICO site._
```{r, echo=FALSE, message=FALSE}
bs_BICO_location <- here::here("results", "BICO_bs_lm.png")

include_graphics(bs_BICO_location)
```

_Supplementary Figure 4: Cook Distance plot for MIDO Biological Score data. Plots show no score >1, suggesting no outliers in the data._
```{r, echo=FALSE, message=FALSE}
MIDO_BS_OUT_location <- here::here("results", "MIDO_BS_OUT.png")

include_graphics(MIDO_BS_OUT_location)
```

_Supplementary Figure 5: Cook Distance plot for MIDO E. coli cfu data. Plots show no score >1, suggesting no outliers in the data._
```{r, echo=FALSE, message=FALSE}
MIDO_EC_OUT_location <- here::here("results", "MIDO_EC_OUT.png")

include_graphics(MIDO_EC_OUT_location)
```

_Supplementary Figure 6: Cook Distance plot for NORO Biological Score data. Only plot showing score >1 is biological score as a function of Nitrate (NO3) concentration. All other plots have a score <1, suggesting no outliers in the data._
```{r, echo=FALSE, message=FALSE}
NORO_BS_OUT_location <- here::here("results", "NORO_BS_OUT.png")

include_graphics(NORO_BS_OUT_location)
```

_Supplementary Figure 7: Cook Distance plot for NORO E. coli cfu data. Only plot showing score >1 is E. coli cfu as a function of Nitrate (NO3) concentration. All other plots have a score <1, suggesting no outliers in the data._
```{r, echo=FALSE, message=FALSE}
NORO_EC_OUT_location <- here::here("results", "NORO_EC_OUT.png")

include_graphics(NORO_EC_OUT_location)
```

_Supplementary Figure 8: Cook Distance plot for BICO Biological Score data. Plots show no score >1, suggesting no outliers in the data._
```{r, echo=FALSE, message=FALSE}
BICO_BS_OUT_location <- here::here("results", "BICO_BS_OUT.png")

include_graphics(BICO_BS_OUT_location)
```

_Supplementary Figure 9: Cook Distance plot for BICO E. coli cfu data. Only plot showing score >1 is E. coli cfu as a function of Turbidity. All other plots have a score <1, suggesting no outliers in the data._
```{r, echo=FALSE, message=FALSE}
BICO_EC_OUT_location <- here::here("results", "BICO_EC_OUT.png")

include_graphics(BICO_EC_OUT_location)
```

### 2021-11-29 products/style_template.Rmd
Include References & Formatting - CC

--- 2021-11-12 products/manuscript/Manuscript.Rmd
@@
 ---
-title: "Relationship Between Common Stream Health Indicators"
-subtitle: ""
-author: Carter Coleman
-date: "`r Sys.Date()`"
-site: bookdown::bookdown_site
-documentclass: article
-output:
-  bookdown::word_document2: 
-    toc: false
-    number_sections: true
-  bookdown::html_document2: 
-    toc: false
-bibliography: ../references.bib
-csl: ../apa.csl
+title: "style_template"
+author: "C. Coleman"
+date: "11/28/2021"
+output: word_document
 ---
 
-# Summary/Abstract
-
-</br>
+```{r, echo=FALSE, results='hide', message=FALSE}
+library(knitr)    # For knitting document and include_graphics function
+library(ggplot2)  # For plotting
+library(png)      # For grabbing the dimensions of png files
+```
 
@@
 
-_When monitoring freshwater streams, stream health is typically measured by its physical characteristics (chemistry, surrounding land use area, channelization, etc.); Microbial (fecal indicator), and ecology (diversity of indicator macroinvertebrates).
-However, these categories are usually analyzed separately to assess stream health.
-As such, my question is how well are these parameters correlated with each other.
-Specifically, can you use one of the indicators mentioned (physical, microbial, and ecological) to describe both of the others to make a statement about overall stream health?_
+_The Earth's surface contains a relative proportion of approximately 3% freshwater to 97% salt water.
+As such, that three percent is supposedly meant to sustain a growing global population of 9 billion people.
+In a world where water scarcity is more and more common due to increasingly warmer climate (@Winter2017), our freshwater reserves are also threatened by pollution by urbanization.
+Urbanization surrounding a freshwater body creates a watershed dominated by impervious surface in the form of concrete and asphalt (@Bouizrou2021).
+Studies have shown that there is a significant negative relationship between impervious surface coverage and water quality (@Birawat2021).
+This is most likely due to accumulation and concentration of run-off pollutants from an entire watershed into a single drainage basin (i.e.
+freshwater body)(@Klein1979; @Wolfram2021).
+Because the urban world continues to grow with population, it is likely that we will see a proportional decrease in water quality in our freshwater bodies as a result (@Wolfram2021).
+Therefore, a rapid and consistent monitoring system needs to be in place to effectively allocate resources to remediating body of water in the midst of being polluted._
 
-</br>
+_Currently, stream health and water quality are typically measured by their physical characteristics (chemistry, surrounding land use area, channelization, etc.); Microbial (fecal indicator), and ecology (diversity of indicator macroinvertebrates) (@Clapcott2012; @Flinders2008; @Alvareda2020).
+Typically, these measurements include conductivity (measure of dissolved ions), dissolved oxygen, temperature, pH, nitrate and phosphorus concentration, turbidity (measure of suspended particulate matter in the water column), fecal indicator bacteria (measure of waste pollution), and benthic macroinvertebrate diversity indices (measure of impervious surface run-off on sensitive aquatic taxa) (@Clapcott2012; @Flinders2008; @Alvareda2020).
+While Biological Indicatior Score is used to measure the combination of stream chemical and physical properties (@Flinders2008), it is hard to evaluate whether urbanization affects watersheds and water quality uniformly (@Wu2020).
+Evaluating how uniform the impact of urbanization is on different watersheds would allow for a cookie-cutter monitoring and remediation method, which would allow implementation to be fairly easy for environmental managers.
+On the other hand, if the impact of urbanization on freshwater bodies is not monolithic, we could be spending a lot of time and effort on monitoring programs that are not effective.
+Currently, citizen science groups, like Adopt-a-Stream in Georgia, are collecting data on many watersheds throughout the world in an effort to consistently monitor water quality in a community.
+However, studies have shown conflicting results on what variables like conductivity, turbidity, e.
+coli cfu (fecal coliform indicator), nutrient (nitrogen and phosphorus) concentration, and biodiversity indicator score actually say about water quality (@Boulton1999).
+Therefore, I aim to answer how well are these parameters correlated with each other in the context of the Adopt-a-Stream method.
+Specifically, can you use one of the indicators mentioned (physical, microbial, and ecological) to describe the others to make a statement about overall stream health?_
 
-_I would expect that because the all three are correlated with urbanization/anthropocentric impact, a negative impact in one indicator category will indicate a correlated negative impact in the other categories.
+_I would expect that because the all three are correlated with urbanization/anthropocentric impact, a negative impact in one indicator category will indicate a correlated negative impact to stream health in the other categories.
 The only caveat is that because level of noise in stream health data, significant correlations may be difficult to find._
@@
 
+</br>
+
+_Figure 1: Plots of E. coli cfu versus conductivity, nitrate concentration, and turbidity for the MIDO site._
+```{r, echo=FALSE, message=FALSE}
+ecfu_MIDO_location <- here::here("results", "MIDO_ecfu_lm.png")
+
+include_graphics(ecfu_MIDO_location)
+```
+
+</br>
+
+_Figure 2: Plots of E. coli cfu versus conductivity, nitrate concentration, and turbidity for the NORO site._
+```{r, echo=FALSE, message=FALSE}
+ecfu_NORO_location <- here::here("results", "NORO_ecfu_lm.png")
+
+include_graphics(ecfu_NORO_location)
+```
+
+</br>
+
+_Figure 3: Plots of E. coli cfu versus conductivity, nitrate concentration, and turbidity for the BICO site._
+```{r, echo=FALSE, message=FALSE}
+ecfu_BICO_location <- here::here("results", "BICO_ecfu_lm.png")
+
+include_graphics(ecfu_BICO_location)
+```
+
 ## Full analysis
 
-_At this point,the main take-aways from data analysis thus far is that correlation of stream health indicators showed that most interactions were non-significant, with the exception of E.
+</br>
+
+_At this point, the main take-aways from data analysis thus far is that correlation of stream health indicators showed that most interactions were non-significant, with the exception of E.
 coli (cfu) and turbidity.
@@
 
-_The null linear models for biological score and E.
-coli cfu produced a RMSE = 6.63, with a standard deviation of 0.13 and RMSE = 0.52, with a standard deviation of 0.014, respectively.
-Results from the biological score LASSO machine learning model indicated that model 47 is the best performing models (RMSE = 6.50; STE = 0.16).
-However, based on comparative RMSE scores, model 47 does not really perform any better than the null model.
-This indicates that it is a bad fit to the MIDO data.
-Results from the E.
-coli cfu LASSO machine learning model indicated that model 41 is the best performing models (RMSE = 0.50; STE = 0.012).
-However, based on comparative RMSE scores, model 41 does not really perform any better than the null model.
-This indicates that it is a bad fit to the MIDO data._
+```{r, echo=FALSE, message=FALSE}
+MIDO_ecfu_con_fit_location <- here::here("results", "MIDO_ecfu_con_fit.png")
 
+include_graphics(MIDO_ecfu_con_fit_location)
+```
 </br>
-
-# Discussion
-
-</br>
-
-_Based on the linear modeling exploratory analysis and LASSO modeling for best combination of predictors, the results indicate that combinations of the variables biological score, E.
-coli cfu, nitrate concentration, conductivity, and turbidity are not able to predict values for biological score or E.
-coli cfu.
-This indicates that one cannot predict one measure of stream health from another.
-This is seen by a lack of consistent significant correlations between each stream health predictor and either biological score or E.
-coli cfu as well as LASSO modeling to find the ideal combination of stream health variables to predict biological score or E.
-coli cfu.
-Given these results, it is likely that the data was biased due to the  inconsistent and opportunistic sampling efforts done by citizen science [find citation for lack of reliability for citizen science data].
-When examining the data set, many values were observed to be missing, with a disproportionate amount of extent observations in what I suspect to be the more easily collected stream health variables.
-For example, significantly more NA observations were included in the biological score column relative to variables like turbidty and conductivity.
-This is because both turbidity and conductivity can be measured in situ with a meter in a matter of minutes, while biological score is calculated following collection and identification of macroinvertebrate families from samples [find reference for biological score assessment methodology].
-Therefore, biological score is much harder to measure.
-Additionally, E.
-coli cfu is processed by culturing E.
-coli from samples and generally requires a trained scientist to do [find reference for E.
-coli cfu qunatification methods].
-While confidence in values may be greater in this case, there is know a shortage of labor due to need for a specialized skill set to process and quantify E.
-coli cfu values._
-
-</br>
-
-## Conclusions
-
-</br>
-
-_Based the results of this study, I would recommend that existing methodology for UOWN stream health characterization and monitoring be re-evaluated.
-The issue being that there seems to be a disparity in sampling effort, both in quality of samples and time spent sampling, for each different stream health metric.
-The result being that the term "healthy stream" does not have a unified definition in terms of the metrics used to quantify stream health.
-I would suggest future scientific studies be conducted to assess to relationship between stream health metrics in order to be able to predict metrics like biological score and E.
-coli cfu so citizen based monitoring does not have to collect these data while still having easier to measure variables that can be used to describe stream health in terms of biological score and E.
-coli cfu._
-
-# References
+_Table 1_

### 2021-11-29 README.Rmd
Supplementary_Figures - CC
Update Readme - CC

@@
 ---
+
+# *Please run the "MADA_Project" file in this folder to ensure use of the correct working directory.*
 
@@
 
-### Question:
+# Getting Started From Scratch
 
-#### When monitoring freshwater streams, stream health is typically measured by its physical characteristics (chemistry, surrounding land use area, channelization, etc.); Microbial (fecal indicator), and ecology (diversity of indicator macroinvertebrates).
+This file contains four main sub-folders titled "code", "data", "products", and "results".
+The "code" folder contains all the code used to process, explore and analyze raw data from the UOWN Site.
+The "data" folder contains the raw and processed data for this project.
+The processed data was produced using the processing script located in the "data" folder.
+The "results" folder contains all figures and tables created from the data exploration and analysis code, located in the "code" folder.
+Finally, the "product" folder contains the final manuscript file, supplementary file, and all accompanying files used to create a reference library in Bibtex and format the manuscript word document.
+
+# Workflow:
+
+1.
+Once in the R project file located in this file, navigate into the "code" file.
+Here, run the processing script to create the processed data.
+This will now be located in the "data" folder, in the sub folder "processed_data".
+
+2.
+In the "code" file, run the "data_exploration" file to run the introductory linear relationships and create figures showing variable relationships.
+The .RMD file will save figures in the "results folder".
+
+3. Next, run the "analysis_uown" file to run the outlier and MIDO LASSO linear analysis models. Results from this file will be saved in the "results" file.
+
+4. Finally, run the "Manuscript" and "Supplementary_Data" files to produce the final article and supplementary report for the UOWN data analysis project.
+
+# Question:
+
+When monitoring freshwater streams, stream health is typically measured by its physical characteristics (chemistry, surrounding land use area, channelization, etc.); Microbial (fecal indicator), and ecology (diversity of indicator macroinvertebrates).
 However, these categories are usually analyzed separately to assess stream health.
@@
 The only caveat is that because level of noise in stream health data, significant correlations may be difficult to find.
-
-#As much as you know, suggest how you will analyze it. At this stage in the course, we haven’t covered analysis approaches yet, so you can keep things vague and non-technical here.
-
-UOWN has provided a great deal of data dating back to 2003.
-First,I will need to wrangle missing data values that are a result of inconsistent sampling efforts.
-This will make the data easier to work with and narrow down a time frame for the data that is consistent and free of sampling bias.
-Next, I will analyze and compare trends in the change of chemical (conductivity, turbidity,	po4.mgL,	no3.mgL); ecological(Biological Score); and microbial indicator data (e.coli CFU).
-Generally speaking, increases in chemical properties like conductivity, turbidity,	po4.mgL,	no3.mgL and microbial indicator abundance (i.e.
-ecoli CFU) indicate decreased stream health.
-While a decrease in biological score indicates decrease in stream health.
-Next, since biological scores are used as a measure of broader ecological health, I will run t-tests on if the other factors significantly effect biological score.
-Finally, I will test the interaction of all three via multivariate analysis.

### 2021-12-02 Project_Review_DDobash_to_CColeman.Rmd
Peer Review- Dawson Dobash to Carter Coleman

---
title: Project Review- Dawson Dobash for Carter Coleman
date: "`r file.mtime(knitr::current_input())`"
#bibliography: ../media/references.bib
output: 
  html_document:
    toc_depth: 3
    number_sections: true
---

# Overview

Title of project: "Relationship Between Common Stream Health Indicators"

Name of project author(s): Carter Coleman

Name of project reviewer: Dawson Dobash

# Specific project content evaluation
Evaluate the different parts of the project by filling in the sections below.

## Background, Context and Motivation
How well is the context of the project described? Is a comprehensive background, including summary of previous/related work given? Is the project well placed into the context of existing work (including proper referencing of existing work). Is it clear why the project was undertaken and what new information it hopes to provide?

### Feedback and Comments

The context of the project is described very well with a lot of background and understanding of what this project is going to be about and what impact it will have. Many related work examples are presented and has a place in this project. It is clear why the project was undertaken. I would say just go back through the introduction as some sentences could be more clear.

### Summary assessment (PICK ONE, DELETE THE OTHERS)
* strong contextualization and motivation

## Question description
How well and clear are the question(s)/hypotheses the project aims to address described? Is it clear how the questions relate to the data?

### Feedback and Comments

Although you mention the question within the introduction section, I would suggest giving  the question its own small section so for those reader that just want to know what you are trying to do can go to that question section. The question itself is clearly related to the data and the information presented within the introduction of this project.

### Summary assessment
* question/hypotheses somewhat explained

## Data description
How well is the data overall described? Is the source provided? Is a codebook or other meta-information available that makes it clear what the data is?

### Feedback and Comments

You gave a list of variables you want to use, but maybe add some type of description for each as the reader will not know what some of these are or how it is measured such as the biological score. I would also provide a link to the website for fast access to the data. Your general desription within the data aquisition is good. I would just add what I mentioned above.

### Summary assessment
* source and overall structure of data somewhat explained

## Data wrangling and exploratory analysis
How well is the data cleaned/processed and explored? Are all steps reasonable and well explained? Are alternatives discussed and considered? Are meaningful exploratory results shown (e.g. in the supplementary materials)?

### Feedback and Comments

You did a lot of cleaning to your data and applying different techniques due to the amount of missing values you had. I like how you had the steps on what you did for cleaning the data as well. For the exploratory results, I would ask Dr. Handel but I dont know if referencing to specific supplement figures is okay in the manuscript or not. Maybe if you want to reference them and if needed to talk about, I would suggest putting the figures in the manuscript. I would also try to reference the specific tables and figures from your manuscript when talking about them in paragraphs such as "In Table/Figure [blank], ..." or "According to Table/Figure [blank],..". I would also, within your supplementary file, describe some of your results. It looks like all of the exploratory results are meaningful for this project.

### Summary assessment
* some weaknesses in wrangling and exploratory component

## Appropriateness of Analysis
Were the analysis methods appropriate for the data? Was the analysis done properly? Were different components of the analysis (e.g. performance measure, variable selection, data pre-processing, model evaluation) done in the best way possible and explained well?

### Feedback and Comments

I think your analyses were appropriate for the data and it was done properly from what I can tell. You included variable selection and other components to your analysis. All of the analyses were explained well and you explained what you did and why you did them.

### Summary assessment
* strong and reasonable analysis

## Presentation
How well are results presented? Are tables and figures easy to read and understand? Are the main figures/tables publication level quality?

### Feedback and Comments

I would suggest having a separate section for the main analysis and exploratory analysis. I think overall the figures and tables look okay and good enough. It could be improved to and presented in a more professional way however. For example figures 4 and 5 hae a mixture of titles. One is from the coding while the other is from the R markdown. I would choose one or the other.

### Summary assessment
* results are very well presented

## Discussion/Conclusions
Are the study findings properly discussed? Are strengths and limitations acknowledged? Are findings interpreted properly?

### Feedback and Comments

I just want to give you a heads up that you gave yourself if you forgot. Within the discussion section you told yourself "[find citation for lack of reliability for citizen science data]." You have this I believe 3 times? I just dont want you to forget. Overall though, the discussion and conclusion looks good and you explained the limitations to the study. Were there any strengths to this study? If so, I would include them.

### Summary assessment
* strong, complete and clear discussion

## Further comments

I would go through your writing within the manuscript as some grammatical errors are present. Also this is just a suggestion like all of my comments but one thing that kind of confused me was within you methods section, you a a paragraph un the methods title but then within 3.3 and 3.4 you have more methods. I would try to combine the paragraph right under methods to the other two sections of 3.3 and 3.4 because I think what your vision was is making that a general description of the methods then going into more detail in 3.3 and 3.4 but for me it was kind of confusing especially when it is split up by the data import and cleaning section.

# Overall project content evaluation
Evaluate overall features of the project  by filling in the sections below.

## Structure
Is the project well structured? Are files in well labeled folders? Do files have reasonable names? Are all "junk" files not needed for analysis/reproduction removed? By just looking at files and folders, can you get an idea of how things fit together?

### Feedback and Comments

Overall the project is structured well with very explicit read me files. The names of the folders are reasonable. The only thing I can think of is maybe putting the data exploration file within the analysis code folder or its separate. I think if you want to keep this in this file, maybe rename it to something like processing and exploration? Overall the read me files makes sure you know where everything is. All junk files are deleted.

### Summary assessment
* well structured

## Documentation 
How well is the project documented? Are you able to understand each step of the whole analysis, each decision that was made, and each line of code? Is enough information provided as comments in code or as part of Rmd files?

### Feedback and Comments

I am going to do this by files. First, within your processing script, comments are very thorough and you explained why you did what you did. For the data exploration file, you did a good job commenting on what each block of code is doing and when you run a model, you added some statistics below it. Finally for the analysis uown file, if I want to be stingy on this, the only thing I would possibly add some small commentary about some results when going along with the code not just the end of a part particularly part 1. But overall analysis uown file is very well commented.

### Summary assessment
* fully and well documented

## Reproducibility
Are all results fully reproducible? Is documentation provided which clearly explains how to reproduce things, and does it work without the need for any manual intervention? Are you able to re-run the whole analysis without having to do manual interventions/edits?

### Feedback and Comments

All of the files are reproducible without manual interventions. The read me files are very detailed and gave step by step directions on how to reproduce the results.

### Summary assessment
* fully reproducible without issues

## Thoroughness
How thorough was the overall study? Were alternatives (e.g. different ways of processing the data or different models) considered? Were alternatives discussed? Were the questions/hypotheses fully and thoroughly addressed?

### Feedback and Comments

Knowing that you had to do separate analyses, you did a good thorough job of making sure each model/analysis was worked out all the way. The only thing is you only used LASSO for the machine learning technique but with the amount of work you already put in and your explanation of the LASSO model is addressing the question well. You did mention that you wanted to look at outliers based on your data and the exploratory analysis.

### Summary assessment
* strong level of thorougness

## Further comments

I would skim through some of your writing as within some files, I see some misspelled words here and there. Also just a comment on your read me files, those are probably the best I have seen. They are very detailed and no questions are asked on what to do after reading those. Great Job overall. Looks like you put in a lot of work!
